---
title: "FastQC Summary Report"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

<style type="text/css">
.main-container {
  max-width: 90%;
  margin-left: auto;
  margin-right: auto;
}
</style>

 
<style type="text/css">
  div.datatables { height: auto !important;}
</style>
  
  
```{r functions, include=FALSE}
library(knitr)
library(reshape2)
library(ggplot2)
library(data.table)
library(DT)
library(RCurl)
library(htmltools)

knitr::opts_chunk$set(echo = TRUE)

addLinkTag<-function(text,link) {
  result<-paste0("<a href='",link,"' target='_blank'>",text,"</a>")
  return(result)
}

figRef <- local({
  tag <- numeric()
  created <- logical()
  used <- logical()
  function(label, caption, prefix = options("figcap.prefix"), 
           sep = options("figcap.sep"), prefix.highlight = options("figcap.prefix.highlight"), trunk.eval=TRUE) {
    if(trunk.eval){ 
      i <- which(names(tag) == label)
      if (length(i) == 0) {
        i <- length(tag) + 1
        tag <<- c(tag, i)
        names(tag)[length(tag)] <<- label
        used <<- c(used, FALSE)
        names(used)[length(used)] <<- label
        created <<- c(created, FALSE)
        names(created)[length(created)] <<- label
      }
      if (!missing(caption)) {
        created[label] <<- TRUE
        paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, 
               " ", caption)
      } else {
        used[label] <<- TRUE
        paste(prefix, tag[label])
      }
    }
  }
})
options(figcap.prefix = "Figure", figcap.sep = ":", figcap.prefix.highlight = "**")


tabRef <- local({
  tag <- numeric()
  created <- logical()
  used <- logical()
  function(label, caption, prefix = options("tabcap.prefix"), 
           sep = options("tabcap.sep"), prefix.highlight = options("tabcap.prefix.highlight"), trunk.eval=TRUE) {
    if(trunk.eval){
      i <- which(names(tag) == label)
      if (length(i) == 0) {
        i <- length(tag) + 1
        tag <<- c(tag, i)
        names(tag)[length(tag)] <<- label
        used <<- c(used, FALSE)
        names(used)[length(used)] <<- label
        created <<- c(created, FALSE)
        names(created)[length(created)] <<- label
      }
      if (!missing(caption)) {
        created[label] <<- TRUE
        paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, 
              " ", caption)
      } else {
        used[label] <<- TRUE
        paste(prefix, tag[label])
      }
    }
  }
})
options(tabcap.prefix = "Table", tabcap.sep = ":", tabcap.prefix.highlight = "**")

check_and_include_graphics<-function(graphicFile) {
  if (file.exists(graphicFile)) {
    include_graphics(graphicFile)
  }
}

```

```{r setup, include=FALSE}
files<-read.table("fileList1.txt", header=FALSE, as.is=TRUE)

getFigure<-function(filename, fileKey, fileDescription){
  return(paste0("```{r,echo=FALSE,results='asis', fig.align='center', fig.cap=figRef('", fileKey, "', '", fileDescription, "')}\n",
"  check_and_include_graphics(\"", filename, "\")\n```\n"))
}

figureFile<-"figures.Rmd"
cnames=c("per_base_quality", "per_base_sequence_content", "per_sequence_quality", "per_sequence_gc_content", "per_tile_quality", "adapter_content")
figureRmd<-function(files, section){
  dnames<-unique(files$V2)
  result<-""
  i<-1
  for(i in c(1:length(dnames))){
    dname = dnames[i]
    result<-paste0(result, "\n# ", dname, "\n\n")

    dfiles<-files[files$V2 == dname,,drop=F]
    j<-1
    for(j in c(1:nrow(dfiles))){
      dfile = dfiles$V1[j]
      folder = dirname(dfile)
      fname = basename(folder)
      fname = gsub("_fastqc", "", fname)
      result<-paste0(result, "\n## ", fname, "\n\n")

      imagefolder = paste0(folder, "/Images/")
      images<-paste0(imagefolder, cnames, ".png")
      images<-images[file.exists(images)]

      result<-paste0(result, "```{r,echo=FALSE,results='asis', fig.show='hold',out.width=\"33%\", fig.align='center', fig.cap=figRef('", fname, "', '", fname, "')}\n")
      result<-paste0(result, "  include_graphics(c(\"", paste0(images, "",  collapse = "\", \""), "\")) \n```\n\n")
    }
  }
  return(result)
}
cat(figureRmd(files, "#"), file=figureFile)

```

```{r, child=figureFile} 
```
