---
title: "scRNA Sequencing Data Analysis Report"
date: "`r format(Sys.time())`"
output: 
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    thumbnails: false
    lightbox: false    
    toc_depth: 3
    number_sections: true
---

<style type="text/css">
#content {
    max-width: 1500px !important;
    margin-left: 300px !important;
}
#table-of-contents {
    width: 300px !important;
}
.main-container {
  max-width: 90%;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 120),
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)

get_cluster_matrix<-function(counts, clusters){
  cluster<-levels(clusters)[1]
  csums=lapply(levels(clusters), function(cluster){
    #cat(cluster, "\n")
    cells=names(clusters)[clusters==cluster]
    subcounts=counts[,cells]
    #cat(ncol(subcounts), "\n")
    Matrix::rowSums(subcounts)
  })
  
  result=do.call(cbind, csums)
  colnames(result)=levels(clusters)
  
  gcount=Matrix::rowSums(result)
  result=result[gcount > 0,]
  return(result)
}

add_cluster<-function(object, new.cluster.name, new.cluster.ids){
  seurat_clusters<-object[["seurat_clusters"]]$seurat_clusters
  names(new.cluster.ids) <- levels(seurat_clusters)

  new.cluster.values<-plyr::mapvalues(x = seurat_clusters, from = levels(seurat_clusters), to = new.cluster.ids)
  names(new.cluster.values)<-names(seurat_clusters)
  
  object[[new.cluster.name]]<-new.cluster.values
  object
}

run_cluster<-function(object, Remove_Mt_rRNA, rRNApattern, Mtpattern, pca_dims, by_sctransform, min.pct = 0.5, logfc.threshold = 0.6){
  if (by_sctransform) {
    object <- RunPCA(object = object, verbose=FALSE)
  }else{
    if (Remove_Mt_rRNA) {
      rRNA.genes <- grep(pattern = rRNApattern,  rownames(object), value = TRUE)
      Mt.genes<- grep (pattern= Mtpattern,rownames(object), value=TRUE )
      var.genes <- dplyr::setdiff(VariableFeatures(object), c(rRNA.genes,Mt.genes))
    } else {
      var.genes <- VariableFeatures(object)
    }
    object <- RunPCA(object = object, features = var.genes, verbose=FALSE)
  }
  object <- RunUMAP(object = object, dims=pca_dims, verbose = FALSE)
  object <- FindNeighbors(object = object, dims=pca_dims, verbose=FALSE)
  object <- FindClusters(object=object, verbose=FALSE, random.seed=random.seed, resolution=resolution)

  if (by_sctransform) {
    markers <- FindAllMarkers(object, only.pos = TRUE, min.pct = min.pct, logfc.threshold = logfc.threshold)
  }else{
    markers <- FindAllMarkers(object, features=var.genes, only.pos = TRUE, min.pct = min.pct, logfc.threshold = logfc.threshold)
  }
  markers <- markers[markers$p_val_adj < 0.01,]
  return(list(object=object, markers=markers))
#  return(list(object=object, markers=NULL))
}

ORA_celltype<-function(medianexp,cellType,weight){
  ORA_result<-matrix(NA, nrow=length(cellType),ncol=dim(medianexp)[2])
  CTA_result<-matrix(0,nrow=length(cellType),ncol=dim(medianexp)[2])
  exp_z<-scale(medianexp)
  genenames<-rownames(medianexp)   
  for (j in 1: dim(medianexp)[2]){
    clusterexp<-medianexp[,j] 
    clusterexp_z<-exp_z[,j]
    for (i in 1:length(cellType)){
      
      ct_exp<-length(intersect(genenames[clusterexp>0],cellType[[i]]))
      ct_not_exp<-length(cellType[[i]])-ct_exp
      exp_not_ct<-sum(clusterexp>0)-ct_exp
      not_exp_not_ct<-length(clusterexp)-ct_not_exp 
      cont.table<-matrix(c(ct_exp,ct_not_exp,exp_not_ct,not_exp_not_ct),nrow=2)
      ORA_result[i,j]<-fisher.test(cont.table,alternative="greater")$p.value
      ###
      weight_ss<-weight[names(weight)%in%cellType[[i]]]
      ind<-match(names(weight_ss),genenames)
      exp_ss<-clusterexp_z[ind[!is.na(ind)]]
      weight_ss<-weight_ss[!is.na(ind)]
      CTA_result[i,j]<-sum(exp_ss*weight_ss)/(length(exp_ss)^(1/3))
    }
  }
  rownames(ORA_result)<-rownames(CTA_result)<-names(cellType)
  minp_ora_ind<- apply(ORA_result,2,function(x){which.min(x)})
  minp_ora<-apply(ORA_result,2,min)
  names(minp_ora)<-rownames(ORA_result)[minp_ora_ind]
  
  max_cta_ind<- apply(CTA_result,2,function(x){which.max(x)})
  max_cta<-apply(CTA_result,2,max,na.rm=T)
  names(max_cta)<-rownames(CTA_result)[max_cta_ind]
  return(list(ora=ORA_result,cta=CTA_result,min_ora=minp_ora,max_cta=max_cta))
}

get_cta_ora_mat<-function(predict_celltype) {
  cta_index<-apply(predict_celltype$cta,2,function(x){return(order(x,decreasing=T)[1:2])})
  cta_index<-unique(sort(cta_index))
  
  cta_mat<- predict_celltype$cta[cta_index,]
  colnames(cta_mat)<-paste0(names(predict_celltype$max_cta), " : ", c(0:(ncol(cta_mat)-1)))
  
  ora_mat<- predict_celltype$ora[cta_index,]
  ora_mat<--log10(ora_mat)  
  colnames(ora_mat)<-colnames(cta_mat)
  
  return(list(cta_mat=cta_mat, ora_mat=ora_mat))
}

get_cta_combined<-function(obj, predicted){
  cta_table<-data.frame(Cluster=c(1:length(predicted$max_cta))-1, 
                        CellType=names(predicted$max_cta),
                        CtaScore=round(predicted$max_cta * 10) / 10.0,
                        stringsAsFactors = F)
  cta_table$OraPvalue=apply(cta_table, 1, function(x){
    ct=x[2]
    cl=as.numeric(x[1]) + 1
    predicted$ora[ct, cl]
  })

  cluster_sample<-as.data.frame.matrix(table(obj$seurat_clusters, obj$orig.ident))

  nc<-apply(cluster_sample, 2, function(x){
    tc=sum(x)
    perc<-x/tc
    return(round(perc*1000) / 10.0)
  })
  colnames(nc)<-paste0(colnames(nc), "_perc")

  cta_combined<-cbind(cta_table, cluster_sample, nc)

  return(cta_combined)
}

```

```{r, child="Functions.Rmd"} 
```

```{r init}
library(dplyr)
library(Seurat)
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(rmdformats)
library(DT)
library(data.table)
library(digest)
library(heatmap3)
library(cowplot)
library(scales)
library(stringr)
library(htmltools)
require(data.table)

options(future.globals.maxSize= 10779361280)
random.seed=20200107

options_table<-read.table("fileList2.txt", sep="\t", header=F, stringsAsFactors = F)
myoptions<-split(options_table$V1, options_table$V2)

dynamic_options<-options_table[!(options_table$V2 %in% c("details_rmd")),]
optionstr<-paste0(dynamic_options$V1, collapse = "_")

Mtpattern= myoptions$Mtpattern
rRNApattern=myoptions$rRNApattern
Remove_Mt_rRNA= ifelse(myoptions$Remove_Mt_rRNA == "FALSE", FALSE, TRUE)
resolution=as.numeric(myoptions$resolution)
species=myoptions$species
markerfile<-myoptions$markers_file
annotate_tcell<-ifelse(myoptions$annotate_tcell == "0", FALSE, TRUE)
HLA_panglao5_file<-myoptions$HLA_panglao5_file
tcell_markers_file<-myoptions$tcell_markers_file
details_rmd<-myoptions$details_rmd
pool_sample<-ifelse(myoptions$pool_sample == "0", FALSE, TRUE)
batch_for_integration<-ifelse(myoptions$batch_for_integration == "0", FALSE, TRUE)
by_integration<-ifelse(myoptions$by_integration == "0", FALSE, TRUE)
by_sctransform<-ifelse(myoptions$by_sctransform == "0", FALSE, TRUE)

remove_subtype_of=ifelse(annotate_tcell, "T cells", "")

hto_str = ""
hto_data = list()
hto_sample_file<-myoptions$hto_sample_file
has_hto = (hto_sample_file != "")
if (has_hto) {
  hto_samples = read.table(hto_sample_file, sep="\t", header=T, stringsAsFactors=FALSE)
  hto_md5 = list()
  hto_md5[["hto_samples"]] = digest(hto_sample_file, file=TRUE)
  hto_cell_files = read.table("fileList5.txt", sep="\t", stringsAsFactors=FALSE, row.names=2)
  #sample = rownames(hto_cell_files)[1]
  for (sample in rownames(hto_cell_files)){
    cell_file = hto_cell_files[sample, "V1"]
    cell_data = read.csv(cell_file, stringsAsFactors=FALSE, header=TRUE, check.names=F)
    hto_md5[[sample]] = digest(cell_file, file=TRUE)
    cell_data$Sample = ""
    cur_samples = hto_samples[hto_samples$File == sample,]
    for (cidx in c(1:nrow(cur_samples))){
      tagname = cur_samples$Tagname[cidx]
      samplename = cur_samples$Sample[cidx]
      cell_data$Sample[cell_data$HTO == tagname] = samplename
    }
    sample_cell_data = cell_data[cell_data$Sample != "",]
    hto_data[[sample]] = sample_cell_data
  }
  hto_str = paste0(hto_md5, collapse="_")
}
prefix<-myoptions$prefix

hasDetails<-file.exists(details_rmd)
if(hasDetails){
  hasDetails<-file.copy(details_rmd, "Details.Rmd", overwrite=T)
}

nFeature_cutoff_min=as.numeric(myoptions$nFeature_cutoff_min)
nFeature_cutoff_max=as.numeric(myoptions$nFeature_cutoff_max)
nCount_cutoff=as.numeric(myoptions$nCount_cutoff)
mt_cutoff=as.numeric(myoptions$mt_cutoff)
#nCount_sd_cutoff=as.numeric(options$nCount_sd_cutoff)
nCount_sd_cutoff=0

pca_dims<-1:as.numeric(myoptions$pca_dims)

filelist1<-read.table("fileList1.txt", header=F)
filestr<-paste0(filelist1$V1, collapse = "_")

prefixMd5=digest(paste0(optionstr, filestr, hto_str), "md5", serialize = FALSE)

finalList<-list()
finalListFile<-paste0(prefix, ".final.rds")
finalListFileOptionsMd5<-paste0(prefix, ".final.options.md5")

isCalc=TRUE
if(file.exists(finalListFile) && file.exists(finalListFileOptionsMd5)){
  oldMd5=str_trim(readChar(finalListFileOptionsMd5, file.info(finalListFileOptionsMd5)$size))
  isCalc=(oldMd5 != prefixMd5)
}

toMouseGeneSymbol<-function(x){
  result=paste0(toupper(substr(x,1,1)),tolower(substr(x,2,nchar(x))))
  return(result)
}

read_cell_markers_file<-function(panglao5_file, species, remove_subtype_of="", HLA_panglao5_file=""){
  #preparing cell activity database
  marker<-data.frame(fread(panglao5_file))
  if(remove_subtype_of != ""){
    remove_subtype_of<-unlist(strsplit(remove_subtype_of, ","))
    pangdb_ct <- read.table(HLA_panglao5_file,header = T,row.names = 1,sep = "\t",stringsAsFactors = F)
    layer="Layer2"
    removed<-c()
    for(layer in c("Layer1","Layer2", "Layer3")){
      layer_ct<-unique(pangdb_ct[,layer])
      for(rs in remove_subtype_of){
        if(rs %in% removed){
          next
        }
        
        if(rs %in% layer_ct){
          subdb<-rownames(pangdb_ct)[pangdb_ct[,layer]==rs]
          if(rs %in% subdb){ 
            subdb<-subdb[subdb != rs]
            marker<-marker[!(marker$cell.type %in% subdb),]
          }else{
            marker$cell.type[marker$cell.type %in% subdb] = rs
          }
          removed<-c(removed, rs)
        }
      }
    }
  }
  
  hsind<-regexpr(species,marker$species)
  marker_species<-marker[hsind>0 & marker$ubiquitousness.index<0.05,]
  if (species=="Mm") {
    ##change the gene symbol only keep the first letter capitalize
    marker_species$official.gene.symbol<-toMouseGeneSymbol(marker_species$official.gene.symbol)
  }
  if (species=="Hs") {
    marker_species$official.gene.symbol<-toupper(marker_species$official.gene.symbol)
  }
  cellType<-tapply(marker_species$official.gene.symbol,marker_species$cell.type,list)
  freq<-sort((table(unlist(cellType)))/length(cellType))
  weight<-1+sqrt((max(freq)-freq)/(max(freq)-min(freq)))
  return(list(cellType=cellType, weight=weight))
}

get_selfdefinedCelltype <- function(file){
  file.ori <- scan(file,what = "",sep="\n")
  marker.ori <- matrix("undefined",nrow=length(file.ori)/2,ncol = 2)
  for (i in 1:length(file.ori)){if(i%%2==1){marker.ori[(i+1)/2,1]=substr(file.ori[i],2,nchar(file.ori[i]))}else{marker.ori[i/2,2]=file.ori[i]}}
  ref <- as.data.frame(marker.ori[,1])
  ref <- tidyr::separate(ref,col=colnames(ref),into = c("celltype","subtypeOf"),sep = ",")
  rownames(ref) <- ref$celltype
  celltype.ori <- sapply(tapply(marker.ori[,2],marker.ori[,1],list), function(x) strsplit(x,","))

  celltype.tem <- as.data.frame(matrix(unlist(celltype.ori)))
  celltype.tem[,2] <- rep(names(celltype.ori),lengths(celltype.ori))
  celltype.tem <- tidyr::separate(celltype.tem,col=V2,into = c("celltype","subtypeOf"),sep = ",")

  ct <- as.data.frame(matrix("Undefined",nrow = nrow(celltype.tem),ncol = length(unique(celltype.tem$subtypeOf))+2),stringsAsFactors=FALSE)
  ct[,1:2] <- celltype.tem[,1:2]
  for (i in 3:ncol(ct)){
    for (j in 1:nrow(ct)) {if(ct[j,i-1] %in% rownames(ref)){ct[j,i]<-ref[ct[j,i-1],2]}else{ct[j,i]<-ct[j,i-1]}}
  }

  layer <- ncol(ct)
  if(! identical(ct[,ncol(ct)],ct[,ncol(ct)-1])){layer <- ncol(ct)}else{
    for (i in 1:ncol(ct)) {if(identical(ct[,i],ct[,i+1])){layer <- i;break}}
  }

  celltype.trim <- ct[,1:layer]
  for (i in 1:nrow(celltype.trim)){
    tag <- which(celltype.trim[i,]==celltype.trim[i,ncol(celltype.trim)])
    if(length(tag)>1){
      if(ncol(celltype.trim)+1-length(tag)==3){suf <- ct[i,3]}
      else{suf <- ct[i,3:(ncol(celltype.trim)+1-length(tag))]}
      pre <- rep(celltype.trim[i,2],length(tag)-1)
      celltype.trim[i,] <- c(celltype.trim[i,][1:2],pre,suf)
    }
  }

  for (i in 1:ncol(celltype.trim)){if(i == 1){colnames(celltype.trim)[i]="gene"}else{colnames(celltype.trim)[i]=paste0("layer",(ncol(celltype.trim)+1-i),"")}}

  ct.final <- celltype.trim
  for (i in 1:ncol(ct.final)){
    if(i >1){ct.final[,i]=celltype.trim[,ncol(celltype.trim)+2-i];colnames(ct.final)[i] <- colnames(celltype.trim)[ncol(celltype.trim)+2-i]}
  }

  cellType<-tapply(ct.final$gene,ct.final$layer3,list)
  freq<-sort((table(unlist(cellType)))/length(cellType))
  weight<-1+sqrt((max(freq)-freq)/(max(freq)-min(freq)))
  return(list(cellType=cellType, weight=weight))
}


if(isCalc){
  #preparing cell activity database
  cell_activity_database<-read_cell_markers_file(markerfile, species, remove_subtype_of, HLA_panglao5_file)
  finalList$cell_activity_database<-cell_activity_database

  if (annotate_tcell) {
    tcell_database<-get_selfdefinedCelltype(tcell_markers_file)
    finalList$tcell_activity_database<-tcell_database
  }
    
  #read raw count dat
  filelist1<-read.table("fileList1.txt", header=F, stringsAsFactors = F)
  rawobjs = list()
  fidx=1
  for (fidx in c(1:nrow(filelist1))) {
    fileName  = filelist1[fidx, 1]
    fileTitle = filelist1[fidx, 2]
    if(dir.exists(fileName)){
      counts = Read10X(fileName)
    } else {
      counts = Read10X_h5(fileName)
    }
    
    adt.counts<-NULL
    if (is.list(counts)){
      adt.counts<-counts$`Antibody Capture`
      counts<-counts$`Gene Expression` 
    }

    if (species=="Mm") {
      rownames(counts)<-toMouseGeneSymbol(rownames(counts))
    }
    if (species=="Hs") {
      rownames(counts)<-toupper(rownames(counts))
    }
    sobj = CreateSeuratObject(counts = counts, project = fileTitle)
    sobj[["percent.mt"]] <- PercentageFeatureSet(object = sobj, pattern = Mtpattern)
    if (!is.null(adt.counts)){
      mat<-as.matrix(adt.counts)
      rowsum<-apply(mat>0, 1, sum)
      mat<-mat[rowsum > (ncol(mat) / 2),]
      sobj[["ADT"]] <- CreateAssayObject(counts = mat)
    }

    if(has_hto && fileTitle %in% names(hto_data)) {
      cell_data = hto_data[[fileTitle]]
      validobj = subset(sobj, cells=cell_data$X)
      #tagname = unique(cell_data$HTO)[1]
      for (tagname in unique(cell_data$HTO)){
        tagcells = cell_data[cell_data$HTO == tagname,]
        sample = tagcells$Sample[1]
        sample_obj = subset(validobj, cells=tagcells$X)
        sample_obj$orig.ident = sample
        sample_obj<-RenameCells(object=sample_obj, new.names=paste0(sample, "_", colnames(sample_obj)))
        rawobjs[[sample]] = sample_obj
      }
    }else{
      sobj<-RenameCells(object=sobj, new.names=paste0(fileTitle, "_", colnames(sobj)))
      rawobjs[[fileTitle]] = sobj
    }
  }
  
  if(pool_sample){
    if(!file.exists("filelist4.txt")){
      stop("pool_sample_groups/filelist4.txt not defined.")
    }
    #cat("pooling samples ... \n")
    pools = read.table("filelist4.txt", header=F, stringsAsFactors = F)
    poolNames = unique(pools$V2)
    pooledObjs = lapply(poolNames, function(pn){
      curPools<-pools[pools$V2==pn,]
      curListIndecies<-which(unlist(lapply(rawobjs, function(x) x@project.name %in% curPools$V1)))
      curObjs<-rawobjs[curListIndecies]
      if(length(curObjs) == 1){
        curobj = curObjs[[1]]
      }else{
        curobj <- merge(curObjs[[1]], y = unlist(curObjs[2:length(curObjs)]), project = "integrated")
      }
      curobj@project.name=pn
      curobj$orig.ident=rep(pn, length(curobj$orig.ident))
      Idents(curobj)<-"orig.ident"
      return(curobj)
    });
    notPoolIndecies<-which(unlist(lapply(rawobjs, function(x) !(x@project.name %in% pools$V1))))
    if(length(notPoolIndecies) > 0){
      notPoolObjs<-rawobjs[notPoolIndecies]
      rawobjs<-c(pooledObjs, notPoolObjs)
    }else{
      rawobjs<-pooledObjs
    }
    #cat("pooling samples done. \n")
  }
  
  if(length(rawobjs) == 1){
    rawobj <- rawobjs[[1]]
  }else{
    rawobj <- merge(rawobjs[[1]], y = unlist(rawobjs[2:length(rawobjs)]), project = "integrated")
  }

  finalList$rawobj<-rawobj

  #filter cells
  finalList$filter<-list(nFeature_cutoff_min=nFeature_cutoff_min,
                        nFeature_cutoff_max=nFeature_cutoff_max,
                        mt_cutoff=mt_cutoff,
                        nCount_cutoff=nCount_cutoff,
                        nCount_sd_cutoff=nCount_sd_cutoff)
  objs<-lapply(rawobjs, function(x){
    sobj<-subset(x, subset = nFeature_RNA > nFeature_cutoff_min & nFeature_RNA<nFeature_cutoff_max & nCount_RNA > nCount_cutoff & percent.mt < mt_cutoff)

    if(nCount_sd_cutoff > 0){
      nCount_mean = mean(sobj[["nCount_RNA"]])
      nCount_sd = sd(sobj[["nCount_RNA"]])
      nCount_sd_min = nCount_mean - nCount_sd_cutoff * nCount_sd
      nCount_sd_max = nCount_mean + nCount_sd_cutoff * nCount_sd
      sobj<-subset(sobj, subset = nCount_RNA > nCount_sd_min & nCount_RNA < nCount_sd_max)
      finalList$filter$nCount_sd_min = nCount_sd_min
      finalList$filter$nCount_sd_max = nCount_sd_max
    }
    return(sobj)
  })  
  rm(rawobjs)

  if(by_integration){
    if(batch_for_integration){
      if(!file.exists("fileList3.txt")){
        stop("batch_for_integration_groups not defined.")
      }
      pools = read.table("filelist3.txt", header=F, stringsAsFactors = F)
      pools$V2=as.character(pools$V2)
      poolNames = unique(pools$V2)
      pooledObjs = lapply(poolNames, function(pn){
        curPools<-pools[pools$V2==pn,]
        curListIndecies<-which(unlist(lapply(objs, function(x) x@project.name %in% curPools$V1)))
        curObjs<-objs[curListIndecies]
        if(length(curObjs) ==1){
          curobj<-curObjs[[1]]
        }else{
          curobj <- merge(curObjs[[1]], y = unlist(curObjs[2:length(curObjs)]), project = "integrated")
        }
        curobj@project.name=pn
        return(curobj)
      });
      notPoolIndecies<-which(unlist(lapply(objs, function(x) !(x@project.name %in% pools$V1))))
      if(length(notPoolIndecies) > 0){
        notPoolObjs<-objs[notPoolIndecies]
        objs<-c(pooledObjs, notPoolObjs)
      }else{
        objs<-pooledObjs
      }
    }

    if(by_sctransform){
      #perform sctransform
      objs<-lapply(objs, function(x){
        x <- SCTransform(x, verbose = FALSE)
        return(x)
      })  
      
      if(length(objs) == 1){
        obj <- objs[[1]]
      }else{
        all_genes <- Reduce(intersect, lapply(objs, rownames)) 
        obj_features <- SelectIntegrationFeatures(object.list = objs, nfeatures = 3000)
        objs <- PrepSCTIntegration(object.list = objs, anchor.features = obj_features, verbose = FALSE)
        obj_anchors <- FindIntegrationAnchors(object.list = objs, normalization.method = "SCT", anchor.features = obj_features, verbose = FALSE)
        obj <- IntegrateData(anchorset = obj_anchors, normalization.method = "SCT", verbose = FALSE, features.to.integrate=all_genes)      
      }
    }else{
      #perform standard workflow
      objs<-lapply(objs, function(x){
        x <- NormalizeData(x, verbose = FALSE)
        x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000, verbose = FALSE)  
        return(x)
      })  
      
      if(length(objs) == 1){
        obj <- objs[[1]]
      }else{
        all_genes <- Reduce(intersect, lapply(objs, rownames)) 
        obj.anchors <- FindIntegrationAnchors(object.list = objs, dims = 1:20)
        obj <- IntegrateData(anchorset = obj.anchors, dims = 1:20, features.to.integrate=all_genes)    
        obj <- ScaleData(obj, verbose = FALSE)
      }
    }
  }else{
    if(length(objs) == 1){
      obj <- objs[[1]]
    }else{
      obj <- merge(objs[[1]], y = unlist(objs[2:length(objs)]), project = "integrated")
    }
    if(by_sctransform){
      #perform sctransform
      obj <- SCTransform(obj, verbose = FALSE)
    }else{
      obj <- NormalizeData(obj, verbose = FALSE)
      obj <- FindVariableFeatures(obj, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
      all.genes <- rownames(obj)  
      obj <- ScaleData(obj, features = all.genes, verbose = FALSE)
    }
  }
  rm(objs)
  
  obj_markers <- run_cluster(obj, Remove_Mt_rRNA, rRNApattern, Mtpattern, pca_dims, by_sctransform)
  
  finalList$markers<-obj_markers$markers
  obj<-obj_markers$object
  seurat_clusters<-unlist(obj[["seurat_clusters"]])
  seurat_colors<-hue_pal()(length(levels(seurat_clusters)))
  names(seurat_colors)<-levels(seurat_clusters)
  
  finalList$seurat_colors<-seurat_colors
  
  counts=GetAssayData(obj,assay="RNA",slot="counts")
  clusters=obj@active.ident
  sumcounts=get_cluster_matrix(counts, clusters)
  logsumcounts<-log2(sumcounts+1)

  data.quantileAll<-apply(logsumcounts, 2, function(x){quantile(x, 0.75)})

  norm_method=""
  if(any(data.quantileAll == 0)){
    norm_method = ".normByTotal"
    data.all <- apply(logsumcounts, 2, sum)
    data.all<-data.all / median(data.all)
    data.norm <- t(t(logsumcounts) / data.all)
  }else{
    norm_method = ".normByUpQuantile"
    data.quantileAll<-data.quantileAll / median(data.quantileAll)
    data.norm <- t(t(logsumcounts) / data.quantileAll)
  }

  colnames(sumcounts)<-paste0("Cluster", colnames(sumcounts))
  write.csv(sumcounts, file=paste0(prefix, ".cluster.count.csv"))
  
  oldname<-colnames(data.norm)
  colnames(data.norm)<-paste0("Cluster", oldname)
  write.csv(data.norm, file=paste0(prefix, ".cluster", norm_method, ".csv"))
  colnames(data.norm)<-oldname
  
  predict_celltype<-ORA_celltype(data.norm,finalList$cell_activity_database$cellType,finalList$cell_activity_database$weight)
  
  new.cluster.ids<-names(predict_celltype$max_cta)
  names(new.cluster.ids) <- levels(seurat_clusters)

  if (annotate_tcell){
    tcell_clusters<-names(new.cluster.ids)[new.cluster.ids=="T cells"]
    tcell_data.norm<-data.norm[,tcell_clusters,drop=F]
    tcell_predict_celltype<-ORA_celltype(tcell_data.norm,finalList$tcell_activity_database$cellType,finalList$tcell_activity_database$weight)
    tcell_new.cluster.ids<-names(tcell_predict_celltype$max_cta)
    names(tcell_new.cluster.ids) <- tcell_clusters
    new.cluster.ids[tcell_clusters] = tcell_new.cluster.ids

    cta_ora_mat = get_cta_ora_mat(tcell_predict_celltype)
    finalList$tcell_cta_mat=cta_ora_mat$cta_mat
    finalList$tcell_ora_mat=cta_ora_mat$ora_mat
    finalList$tcell_activity_database$predicted<-tcell_predict_celltype
  }

  obj[["cellactivity_clusters"]] <- new.cluster.ids[unlist(obj[["seurat_clusters"]])]
  
  clusterDf<-data.frame(seurat=unlist(obj[["seurat_clusters"]]), cellactivity=unlist(obj[["cellactivity_clusters"]]))
  clusterDf$seurat_colors<-seurat_colors[clusterDf$seurat]
  
  clusterDf$seurat_cellactivity<-paste0(clusterDf$seurat, " : ", clusterDf$cellactivity)
  seurat_cellactivity<-clusterDf$seurat_cellactivity
  
  caCount<-table(clusterDf$cellactivity)
  clusterDf$caCount<-caCount[clusterDf$cellactivity]
  
  clusterDf<-clusterDf[order(-clusterDf$caCount, clusterDf$seurat),]
  seurat_cellactivity<-factor(seurat_cellactivity, levels=unique(clusterDf$seurat_cellactivity))
  seurat_cellactivity_colors<-unique(clusterDf$seurat_colors)
  obj[["seurat_cellactivity_clusters"]] <- seurat_cellactivity
  
  clusters<-data.frame("cell" = c(1:length(obj$seurat_clusters)), "seurat_clusters"=as.numeric(as.character(obj$seurat_clusters)), "cellactivity_clusters"=obj$cellactivity_clusters, "seurat_cellactivity_clusters"=obj$seurat_cellactivity_clusters, stringsAsFactors = F)
  rownames(clusters)<-names(obj$seurat_clusters)
  write.csv(clusters, file=paste0(prefix, ".cluster.csv"))
    
  cta_ora_mat = get_cta_ora_mat(predict_celltype)
  finalList$cta_mat = cta_ora_mat$cta_mat
  finalList$ora_mat = cta_ora_mat$ora_mat
  finalList$cell_activity_database$predicted<-predict_celltype

  finalList$seurat_cellactivity_colors<-seurat_cellactivity_colors
  finalList$obj<-obj
  saveRDS(finalList, file=finalListFile)
  
  fileConn<-file(finalListFileOptionsMd5)
  writeLines(c(prefixMd5), fileConn)
  close(fileConn)
}else{
  finalList<-readRDS(file=finalListFile)
  if(!is.null(finalList$tcell_predict_celltype)){
    finalList$tcell_activity_database$predicted=finalList$tcell_predict_celltype
  }
}
obj<-finalList$obj
seurat_colors<-finalList$seurat_colors
seurat_cellactivity_colors<-finalList$seurat_cellactivity_colors
hasMultipleSamples<-length(unique(obj$orig.ident))>1
hasBiomarker<-!is.null(finalList$markers)

samples<-unique(finalList$rawobj$orig.ident)
if(length(samples) < 3){
  sampleHeight=4
}else{
  sampleHeight=min(10, ceiling(length(samples) / 3) * 4)
}

isLargeDataset=ncol(obj) > 30000
if(isLargeDataset){
  downsample=round(300/length(unique(obj$seurat_clusters))) * 100
  heatmap_obj=subset(obj, downsample=downsample)
}else{
  heatmap_obj=obj
}
```

# Quality control

A few QC metrics [commonly used](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4758103/) by the community include

* The number of unique genes detected in each cell. 
    + Low-quality cells or empty droplets will often have very few genes
    + Cell doublets or multiplets may exhibit an aberrantly high gene count
* Similarly, the total number of molecules detected within a cell (correlates strongly with unique genes)
* The percentage of reads that map to the mitochondrial genome
    + Low-quality / dying cells often exhibit extensive mitochondrial contamination
    + We calculate mitochondrial QC metrics with the `PercentageFeatureSet` function, which calculates the percentage of counts originating from a set of features
    + We use the set of all genes starting with either `mt-` or `MT-` as a set of mitochondrial genes

We visualize QC metrics.

```{r qc2, fig.height=6, fig.width=13, fig.cap=figRef("qc_metrics_2", "Scatter plot of number of feature/percentage of mitochondrial vs number of read count")}
plot1 <- FeatureScatter(object = finalList$rawobj, feature1 = "nCount_RNA", feature2 = "percent.mt") + geom_hline(yintercept = mt_cutoff, color="black")  + geom_vline(xintercept = nCount_cutoff, color="black")
plot2 <- FeatureScatter(object = finalList$rawobj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")  + geom_hline(yintercept = c( nFeature_cutoff_min, nFeature_cutoff_max), color="black")  + geom_vline(xintercept = nCount_cutoff, color="black") 
p<-CombinePlots(plots = list(plot1,plot2))
print(p)
```

```{r qc2_1, fig.height=sampleHeight, fig.width=13, fig.cap=figRef("qc_metrics_2_1", "Density plot of number of read vs percentage of mitochondrial")}
mt<-data.frame(mt=finalList$rawobj$percent.mt, Sample=finalList$rawobj$orig.ident, nFeature=log10(finalList$rawobj$nFeature_RNA), nCount=log10(finalList$rawobj$nCount_RNA))
g1<-ggplot(mt, aes(x=mt,y=nCount) ) +
  geom_bin2d(bins = 70) + 
  scale_fill_continuous(type = "viridis") + 
  geom_vline(xintercept = mt_cutoff, color="red")  + 
  geom_hline(yintercept = log10(nCount_cutoff), color="red") +
  xlab("Percentage of mitochondrial") + ylab("log10(number of read)") +
  facet_wrap(~Sample) + theme_bw() + theme(strip.background = element_rect(colour="black", fill="white"))
print(g1)
```

```{r qc2_2, fig.height=sampleHeight, fig.width=13, fig.cap=figRef("qc_metrics_2_2", "Density plot of number of feature vs percentage of mitochondrial")}
g1<-ggplot(mt, aes(x=mt,y=nFeature) ) +
  geom_bin2d(bins = 70) + 
  scale_fill_continuous(type = "viridis") + 
  geom_vline(xintercept = mt_cutoff, color="red")  + 
  geom_hline(yintercept = log10(nFeature_cutoff_min), color="red") +
  geom_hline(yintercept = log10(nFeature_cutoff_max), color="red") +
  xlab("Percentage of mitochondrial") + ylab("log10(number of feature)") +
  facet_wrap(~Sample) + theme_bw() + theme(strip.background = element_rect(colour="black", fill="white"))
print(g1)
```

# Data filter and cluster analysis 

We use QC metrics to filter cells.

* We filter cells that have unique feature counts less than `r nFeature_cutoff_min` or larger than `r nFeature_cutoff_max` 
* We filter cells that have read count less than `r nCount_cutoff`
* We filter cells that have >`r mt_cutoff`% mitochondrial counts

```{r data_filter, results="asis"}
rawCells<-data.frame(table(finalList$rawobj$orig.ident))
filteredCells<-data.frame(table(finalList$obj$orig.ident))
qcsummary<-merge(rawCells, filteredCells, by = "Var1")
colnames(qcsummary)<-c("Sample", "RawCell", "ValidCell")
qcsummary$DiscardCell<-qcsummary$RawCell-qcsummary$ValidCell
qcsummary$DiscardRate<-qcsummary$DiscardCell / qcsummary$RawCell
print(kable(qcsummary, caption=tabRef("qcsummary", "Cell count in sample")) %>% kable_styling() %>% htmltools::HTML())
```

## Marker genes for each cluster

```{r markers, results="asis", eval=hasBiomarker}
clusterMarkers<-finalList$markers %>% group_by(cluster)
write.csv(clusterMarkers, file=paste0(prefix, ".allmarkers.csv"), row.names=F, quote = F)

if ("Seurat" %in% names(sessionInfo()$otherPkgs) & grepl("^4",sessionInfo()$otherPkgs$Seurat$Version)) { #Seurat version4
  logFcColName="avg_log2FC"
} else {
  logFcColName="avg_logFC"
}

#top10 <- finalList$markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
top10 <- finalList$markers %>% group_by(cluster) %>% top_n(n = 10, wt = .data[[logFcColName]])
top10marker_file = paste0(prefix, ".top10markers.csv")
write.csv(top10, file=top10marker_file, row.names=F, quote = F)

top5 <- finalList$markers %>% group_by(cluster) %>% top_n(n = 5, wt = .data[[logFcColName]])
print(kable(top5, caption=tabRef("top5", "Top 5 marker genes in each cluster")) %>% kable_styling() %>% htmltools::HTML())
```

# Cell clusters
    
```{r qc1, fig.height=5, fig.width=13, fig.cap=figRef("qc_metrics_1", "Violin plot of percentage of mitochondrial, number of feature and number of read count")}
g<-VlnPlot(object = obj, features = c("percent.mt", "nFeature_RNA", "nCount_RNA"))
print(g)
```

```{r umap_cells, fig.height=8, fig.width=18, fig.cap=figRef("umap_cells", ifelse(hasMultipleSamples, "UMAP of cells (left) and samples (right)", "UMAP of cells"))}
# note that you can set `label = TRUE` or use the LabelClusters function to help label individual clusters
p1<-DimPlot(object = obj, reduction = 'umap', label=TRUE, group.by="seurat_clusters", cols=seurat_colors)
if(hasMultipleSamples){
  p2<-DimPlot(object = obj, reduction = 'umap', label=FALSE, group.by="orig.ident")
  plot_grid(p1,p2)
}else{
  print(p1)
}
```

```{r sample_cells_count, eval=hasMultipleSamples, fig.height=8, fig.width=18, fig.cap=figRef("sample_cells_count", "Percentage of cell from samples in each cluster (total cell in each sample were normalized to equal number)")}
os<-table(obj$orig.ident, obj$seurat_clusters)
osn<-os / rowSums(os) * 5000
osm<-reshape2::melt(osn)
colnames(osm)<-c("Sample", "Cluster", "Cell")
osm$Cluster<-paste0("Cluster ", osm$Cluster)
osm$Cluster<-factor(osm$Cluster, levels=unique(osm$Cluster))
osm_perc = osm %>% dplyr::group_by(Cluster) %>% dplyr::mutate(Percent = Cell/sum(Cell))
g<-ggplot(osm_perc, aes(x="",y=Percent,fill=Sample)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0) + facet_wrap(~Cluster) + theme_void()
print(g)

```

```{r eval=isLargeDataset, comment=''}
cat("There are total ", ncol(obj), " cells, so each cluster was downsampled to ", downsample, " cells for heatmap.")
```

```{r heatmap_samples, fig.height=max(10, length(top10$gene) / 10), fig.width=12, fig.cap=figRef("heatmap_cluster", "Marker genes in cluster")}
g<-DoHeatmap(heatmap_obj, features = top10$gene, group.colors=seurat_colors, angle = 90) + NoLegend()
print(g)

pdf(file=paste0(prefix, ".marker_cluster.heatmap.pdf"), width=12, height=max(10, length(top10$gene) / 10))
g<-DoHeatmap(obj, features = top10$gene, group.colors=seurat_colors, angle = 90) + NoLegend()
print(g)
```

# Predict cell types

## Cell type prediction using gene activity database

### Predicted cell types for each cluster by cell activity analysis

```{r gene_activity_tbl, results="asis"}

cta_combined<-get_cta_combined(obj, finalList$cell_activity_database$predicted)

write.csv(cta_combined, file=paste0(prefix, ".cluster.summary.csv"), row.names=F, quote=F)

print(kable(cta_combined, caption=tabRef("cta_table", "Cell type prediction result")) %>% kable_styling() %>% htmltools::HTML())
```

```{r tcell_gene_activity_tbl, results="asis", eval=annotate_tcell}
cta_combined<-get_cta_combined(obj, finalList$tcell_activity_database$predicted)

write.csv(cta_combined, file=paste0(prefix, ".tcell.cluster.summary.csv"), row.names=F, quote=F)

print(kable(cta_combined, caption=tabRef("tcell_cta_table", "T cell subtype prediction result")) %>% kable_styling() %>% htmltools::HTML())
```

```{r gene_activity_umap, fig.width=18, fig.height=10, results="asis", fig.cap=figRef("umap_cluster_cell", "Predicted cell types")}
p2<-DimPlot(object = obj, reduction = "umap",label=T, group.by="seurat_cellactivity_clusters", cols=seurat_cellactivity_colors)
print(p2)
```

```{r gene_activity_umap_samples, eval=hasMultipleSamples, fig.height=8, fig.width=18, fig.cap=figRef("gene_activity_umap_samples", "Predicted cell types by sample")}
g=DimPlot(obj, reduction = "umap", label=T, group.by="seurat_clusters", cols=seurat_colors, split.by = "orig.ident") + NoLegend()
print(g)
```

```{r activity_by_samples, eval=hasMultipleSamples, fig.height=8, fig.width=18, fig.cap=figRef("activity_by_samples", "Predicted cell types by sample")}

os<-table(obj$orig.ident, obj$cellactivity_clusters)
osn<-os / rowSums(os) * 5000
osm<-reshape2::melt(osn)
colnames(osm)<-c("Sample", "Cluster", "Cell")
# osm_perc = osm %>% dplyr::group_by(Cluster) %>% dplyr::mutate(Percent = Cell/sum(Cell))
# g<-ggplot(osm_perc, aes(x="",y=Percent,fill=Sample)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0) + facet_wrap(~Cluster) + theme_void()
# print(g)

osm_perc = osm %>% dplyr::group_by(Sample) %>% dplyr::mutate(Percent = Cell/sum(Cell))
g<-ggplot(osm_perc, aes(x="",y=Percent,fill=Cluster)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0) + facet_wrap(~Sample) + theme_void()
print(g)
```

```{r gene_activity_heatmap, fig.width=12, fig.height=max(10, length(top10$gene) / 10), results="asis", fig.cap=figRef("cell_activity_heatmap", "Predicted cell type cluster heatmap")}
g=DoHeatmap(heatmap_obj, features = top10$gene, group.by="seurat_cellactivity_clusters", group.colors=seurat_cellactivity_colors, angle=90) + NoLegend()
print(g)


pdf(file=paste0(prefix, ".marker_celltype.heatmap.pdf"), width=12, height=max(10, length(top10$gene) / 10))
g=DoHeatmap(obj, features = top10$gene, group.by="seurat_cellactivity_clusters", group.colors=seurat_cellactivity_colors, angle=90) + NoLegend()
print(g)

```

```{r gene_activity_cta_score_heatmap, fig.width=18, fig.height=10, results="asis", fig.cap=figRef("cell_activity_cta_heatmap", "Predicted cell type activity score")}
heatmap3(finalList$cta_mat, scale="none", margins=c(10,5))
```


```{r gene_activity_tcell_cta_score_heatmap, fig.width=18, fig.height=10, results="asis", eval=annotate_tcell, fig.cap=figRef("tcell_activity_cta_heatmap", "Predicted T cell subtype activity score")}
heatmap3(finalList$tcell_cta_mat, scale="none", margins=c(10,5))
```

```{r, child="Details.Rmd", eval=hasDetails} 
```

# Project options

```{r options, results="asis"}
if(hasDetails){
  file.remove("Details.Rmd")
}
colnames(options_table)<-c("Value", "Option")
options_table<-options_table[,c("Option", "Value")]
options_table<-options_table[!(options_table$Option %in% c("Mtpattern", "rRNApattern")),]
print(kable(options_table, caption=tabRef("options_table", "Options table")) %>% kable_styling() %>% htmltools::HTML())
```
