---
title: "Cell Image Report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: hide
    number_sections: false
params:
  report_csv: null
---

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

if (is.null(params$report_csv) || !nzchar(params$report_csv)) {
  stop("Parameter 'report_csv' is required.")
}

report_csv <- normalizePath(params$report_csv, mustWork = TRUE)
report_dir <- dirname(report_csv)
report_df <- read.csv(
  report_csv,
  stringsAsFactors = FALSE,
  check.names = FALSE,
  colClasses = "character"
)

if (!nrow(report_df)) {
  stop("The report CSV has no rows.")
}

required_columns <- c("cell_group", "cell_id", "cell_image", "cell_width", "cell_height")
missing_columns <- setdiff(required_columns, colnames(report_df))
if (length(missing_columns) > 0) {
  stop(sprintf(
    "Missing required columns in report CSV: %s",
    paste(missing_columns, collapse = ", ")
  ))
}

image_specs <- list(
  list(title = "Cell Images", image_col = "cell_image", width_col = "cell_width", height_col = "cell_height")
)

if ("nucleus_image" %in% colnames(report_df)) {
  nucleus_required <- c("nucleus_width", "nucleus_height")
  missing_nucleus <- setdiff(nucleus_required, colnames(report_df))
  if (length(missing_nucleus) > 0) {
    stop(sprintf(
      "Missing required nucleus columns in report CSV: %s",
      paste(missing_nucleus, collapse = ", ")
    ))
  }

  image_specs[[length(image_specs) + 1]] <- list(
    title = "Nucleus Images",
    image_col = "nucleus_image",
    width_col = "nucleus_width",
    height_col = "nucleus_height"
  )
}

format_cell_id <- function(cell_id) {
  if (is.na(cell_id)) {
    return("")
  }

  cell_text <- as.character(cell_id)
  if (grepl("^[0-9]+\\.0+$", cell_text)) {
    return(sub("\\.0+$", "", cell_text))
  }

  cell_text
}

resolve_image_path <- function(image_path) {
  if (grepl("^(/|[A-Za-z]:[/\\\\])", image_path)) {
    return(image_path)
  }

  file.path(report_dir, image_path)
}

max_dimension <- 0
for (spec in image_specs) {
  max_dimension <- max(
    max_dimension,
    suppressWarnings(max(as.numeric(report_df[[spec$width_col]]), na.rm = TRUE)),
    suppressWarnings(max(as.numeric(report_df[[spec$height_col]]), na.rm = TRUE))
  )
}

if (!is.finite(max_dimension) || max_dimension <= 0) {
  max_dimension <- 1
}

scale_factor <- if (max_dimension > 180) 180 / max_dimension else 1

render_card <- function(cell_group, cell_id, image_path, section_title, image_width, image_height) {
  resolved_path <- resolve_image_path(image_path)
  if (!file.exists(resolved_path)) {
    stop(sprintf("Image file not found: %s", resolved_path))
  }

  display_width <- max(1L, as.integer(round(as.numeric(image_width) * scale_factor)))
  display_height <- max(1L, as.integer(round(as.numeric(image_height) * scale_factor)))
  formatted_cell_id <- format_cell_id(cell_id)
  alt_text <- htmltools::htmlEscape(sprintf("%s %s %s", cell_group, formatted_cell_id, section_title))
  caption <- htmltools::htmlEscape(formatted_cell_id)
  image_uri <- knitr::image_uri(resolved_path)

  sprintf(
    '<figure class="image-card"><img src="%s" alt="%s" width="%d" height="%d" /><figcaption>%s</figcaption></figure>',
    image_uri,
    alt_text,
    display_width,
    display_height,
    caption
  )
}
```

<style type="text/css">
body {
  font-family: Arial, sans-serif;
  margin: 24px;
  color: #1f2933;
}

h1, h2, h3 {
  margin-bottom: 8px;
}

p {
  margin-top: 0;
}

.group-section {
  margin-bottom: 32px;
}

.grid-wrapper {
  overflow-x: auto;
  padding-bottom: 8px;
}

.image-grid {
  display: grid;
  grid-template-columns: repeat(10, max-content);
  gap: 12px;
  align-items: start;
}

.image-card {
  margin: 0;
  text-align: center;
}

.image-card img {
  display: block;
  border: 1px solid #d2d6dc;
  background: #ffffff;
}

.image-card figcaption {
  margin-top: 4px;
  font-size: 12px;
  white-space: nowrap;
}
</style>

All images are embedded directly into this HTML report. Scale factor: `r sprintf("%.3fx", scale_factor)`.

```{r report, results='asis'}
for (cell_group in sort(unique(report_df$cell_group))) {
  group_df <- report_df[report_df$cell_group == cell_group, , drop = FALSE]

  cat("#", cell_group, "\n\n")
  group_df = group_df[order(group_df$cell_id), , drop = FALSE]
  for (spec in image_specs) {
    cat('<div class="grid-wrapper"><div class="image-grid">')
    for (row_idx in seq_len(nrow(group_df))) {
      row <- group_df[row_idx, , drop = FALSE]
      cat(
        render_card(
          cell_group = row[["cell_group"]][[1]],
          cell_id = row[["cell_id"]][[1]],
          image_path = row[[spec$image_col]][[1]],
          section_title = spec$title,
          image_width = row[[spec$width_col]][[1]],
          image_height = row[[spec$height_col]][[1]]
        )
      )
    }

    cat('</div></div>')
  }
}
```
