---
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: hide
    number_sections: yes
---

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(Hmisc)
require(arsenal)
library(gtsummary)
library(flextable)

knitr::opts_chunk$set(
  echo=TRUE, 
  include=TRUE, 
  warning=FALSE, 
  message=FALSE, 
  results="asis"
)

date_str = format(Sys.time(), "%Y%m%d")

#devtools::install_github("jinworks/CellChat")
library(CellChat)

source('reportFunctions.R')
source('scRNA_func.r')
```

```{r}
myoptions=read_file_map('fileList1.cellchat_summary.txt', do_unlist=FALSE)
circle_width=to_numeric(myoptions$circle_width, 5)
circle_height=to_numeric(myoptions$circle_height, 6)
chord_width=to_numeric(myoptions$chord_width, 8)
chord_height=to_numeric(myoptions$chord_height, 8.5)
```

---
title: "`r myoptions$task_name` - CellChat"
author: 
  email: `r myoptions$email`
  affiliation: `r myoptions$affiliation`
---

```{r}
draw_circle<-function(cur_mat, cur_png, groupSize){
  png(cur_png, width=circle_width, height=circle_height, units="in", res=300, bg="white")
  netVisual_circle(cur_mat, 
    vertex.weight = groupSize, 
    weight.scale = T, 
    label.edge= F, 
    title.name = "Interaction weights/strength",
    margin=c(0.2, 0, 0.5, 0))
  ignored=dev.off()
}
```

```{r}
rds_files=read_file_map('fileList2.cellchat_summary.txt', do_unlist=FALSE)
```

```{r}
group_content=""
for(cur_group in names(rds_files)){
  group_content=paste0(group_content, "```{r}\n")
  group_content=paste0(group_content, "cur_group='", cur_group, "'\n")
  group_content=paste0(group_content, "cur_rds='", rds_files[[cur_group]], "'\n")
  group_content=paste0(group_content, "```\n\n")
  group_content=paste0(group_content, "```{r child='cellchat_summary_sub.rmd'}\n")
  group_content=paste0(group_content, "```\n\n")
}
group_rmd="group.rmd"
writeLines(group_content, group_rmd)
```

```{r child=group_rmd}
```

# Save the session information

```{r}
writeLines(capture.output(sessionInfo()), paste0(date_str, '.sessionInfo.txt'))
```
