```{r eval=FALSE, include=FALSE}
cur_comp="HGPS_vs_WT"
```

# Comparison: `r cur_comp`

```{r}
cur_tbl=comparison_tbl |> dplyr::filter(comparison == cur_comp)
ctrl_group=cur_tbl$group[1]
case_group=cur_tbl$group[2]
cur_list=object_list[c(ctrl_group, case_group)]
cur_cellchat <- mergeCellChat(cur_list, add.names = names(cur_list))
color_use=c("#00BFC4", "#F8766D")
#somehow named color_use is not working, have to use unnamed vector.
#names(color_use) = c(ctrl_group, case_group)

if(!dir.exists("details")){
  dir.create("details")
}
file_prefix=paste0("details/", cur_comp)
```

- Control group: `r ctrl_group`
- Case group: `r case_group`

## Number of cells in each cell type

```{r}
ctrl_df=as.data.frame(table(object_list[[ctrl_group]]@meta$labels)) |>
  dplyr::rename(Cell_type=1, Ctrl=2)
colnames(ctrl_df)[2]=ctrl_group

case_df=as.data.frame(table(object_list[[case_group]]@meta$labels)) |>
  dplyr::rename(Cell_type=1, Case=2)
colnames(case_df)[2]=case_group

cell_type_df=merge(ctrl_df, case_df, by="Cell_type", all=TRUE, fill=0) 
cell_type_df=cell_type_df[order(rowSums(cell_type_df[, c(ctrl_group, case_group)]), decreasing = TRUE),]
print_table(cell_type_df, byDT=FALSE, row.names=FALSE)
```

## Total number of interactions and interaction strength

- The number of interactions is a count of the total unique signaling pathways between cell types. 
- The interaction strength is the ligand-receptor communication probability. 

```{r}
gg1 <- compareInteractions(cur_cellchat, color.use=color_use, show.legend = F, group = c(1,2))
gg2 <- compareInteractions(cur_cellchat, color.use=color_use, show.legend = F, group = c(1,2), measure = "weight")
g = gg1 + gg2

interaction_png=paste0(file_prefix, ".interaction_bar.png")
ggsave(interaction_png, g, width=6, height=3, units="in", dpi=300)
include_graphics(interaction_png)
```

## Number of interactions circle plot

The width of edges represent the strength of the communication.

```{r}
weight.max <- getMaxWeight(cur_list, attribute = c("idents","count"))
g1 = function() netVisual_circle(cur_list[[1]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(cur_list)[1]))
g2 = function() netVisual_circle(cur_list[[2]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(cur_list)[2]))

circle_png=paste0(file_prefix, ".circle.png")
png(circle_png, width=16, height=10, units="in", res=300, bg="white")
plot_grid(g1, g2, ncol = 2)
ignored=grDevices::dev.off()
include_graphics(circle_png)
```

## Outgoing and incoming interactions

This scatter plot shows the dominant senders (sources/outgoing) and receivers (targets/incoming) in a 2D space. x-axis and y-axis are 
respectively the total outgoing or incoming communication probability associated with each cell group. Dot size is
proportional to the number of inferred links (both outgoing and incoming) associated with each cell group.

```{r}
num.link <- sapply(cur_list, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
x_range <- c(Inf, -Inf)
y_range <- c(Inf, -Inf)
for (i in 1:length(cur_list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter( cur_list[[i]], 
                                                title = names(cur_list)[i], 
                                                weight.MinMax = weight.MinMax) + 
              theme(aspect.ratio = 1)
  built <- ggplot2::ggplot_build(gg[[i]])
  x_tmp <- built$layout$panel_params[[1]]$x.range
  y_tmp <- built$layout$panel_params[[1]]$y.range
  x_range <- c(min(x_range[1], x_tmp[1]), max(x_range[2], x_tmp[2]))
  y_range <- c(min(y_range[1], y_tmp[1]), max(y_range[2], y_tmp[2]))
}
gg <- lapply(gg, function(p) p + coord_cartesian(xlim = x_range, ylim = y_range))
g = patchwork::wrap_plots(plots = gg)
dot_png=paste0(file_prefix, ".dot_plot.png")
ggsave(dot_png, g, width=10, height=5, units="in", dpi=300, bg="white")
include_graphics(dot_png)
```

## Differential number of interactions and interaction strength in cell types

The differential number of interactions or interaction strength in the cell-cell communication 
network between two groups. Red and blue represent increased or decreased signaling in `r case_group` 
comparing to `r ctrl_group`.   

### Circle plot

```{r}
g1 = function() netVisual_diffInteraction(cur_cellchat, weight.scale = T, title.name = "Differential No. of interactions")
g2 = function() netVisual_diffInteraction(cur_cellchat, weight.scale = T, measure = "weight", title.name = "Differential interaction strength")

diff_interactions_png=paste0(file_prefix, ".diff_interactions.png")
png(diff_interactions_png, width=16, height=10, units="in", res=300, bg="white")
plot_grid(g1, g2, ncol = 2)
ignored=grDevices::dev.off()
include_graphics(diff_interactions_png)
```

### Heatmap plot

This heatmap shows differential number of interactions and interaction strength in the cell-cell
communication network between two groups. Red and blue represent increased or decreased signaling 
in `r case_group` comparing to `r ctrl_group`.;

```{r}
gg1 <- netVisual_heatmap(cur_cellchat)
gg2 <- netVisual_heatmap(cur_cellchat, measure = "weight")
g = gg1 + gg2

interaction_heatmap_png=paste0(file_prefix, ".interaction_heatmap.png")
png(interaction_heatmap_png, width=10, height=5, units="in", res=300, bg="white")
print(g)
ignored=dev.off()
include_graphics(interaction_heatmap_png)
```

## Altered pathway signaling

Rank signaling networks based on the information flow or the number of
interactions.

```{r}
gg1 <- rankNet(cur_cellchat, mode = "comparison", color.use=color_use, measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE)
altered_pathway_png=paste0(file_prefix, ".altered_pathway.png")
ggsave(altered_pathway_png, gg1, width=6, height=14, units="in", dpi=300)
include_graphics(altered_pathway_png)
```

## Information flow in each cell population

CellChat can also compare the outgoing (or incoming) signaling pattern between two groups, 
allowing to identify signaling pathways/ligand-receptors that exhibit different signaling patterns. 
The heatmap shows the contribution of signals to cell groups in terms of outgoing or incoming signaling. The colorbar represents the relative 
signaling strength of a signaling pathway across cell groups. The top colored bar plot shows the total signaling strength of a cell group by summarizing all signaling pathways 
displayed in the heatmap. The right grey bar plot shows the total signaling strength of a signaling pathway by 
summarizing all cell groups displayed in the heatmap.   

### Outgoing signaling

```{r}
pathway.union <- union(cur_list[[1]]@netP$pathways, cur_list[[2]]@netP$pathways)
#somehow, using height=30 can show all the labels, when creating png file, have to use height=15 to remove the extra white space.
height=30
width=4.5
ht1 = netAnalysis_signalingRole_heatmap(cur_list[[1]], pattern = "outgoing", signaling = pathway.union, title = names(cur_list)[1], width = width, height = height)
ht2 = netAnalysis_signalingRole_heatmap(cur_list[[2]], pattern = "outgoing", signaling = pathway.union, title = names(cur_list)[2], width = width, height = height)
role_heatmap_png=paste0(file_prefix, ".role_heatmap_outgoing.png")
png(role_heatmap_png, width=width*2, height=15, units="in", res=300, bg="white")
draw(ht1 + ht2, ht_gap = unit(2, "cm"))
ignored=dev.off()
include_graphics(role_heatmap_png)
```

### Incoming signaling

```{r}
pathway.union <- union(cur_list[[1]]@netP$pathways, cur_list[[2]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(cur_list[[1]], 
                                        pattern = "incoming", 
                                        signaling = pathway.union, 
                                        title = names(cur_list)[1], 
                                        width = width, 
                                        height = height)
ht2 = netAnalysis_signalingRole_heatmap(cur_list[[2]], 
                                        pattern = "incoming", 
                                        signaling = pathway.union, 
                                        title = names(cur_list)[2], 
                                        width = width, 
                                        height = height)
role_heatmap_png=paste0(file_prefix, ".role_heatmap_incoming.png")
png(role_heatmap_png, width=width*2, height=15, units="in", res=300, bg="white")
draw(ht1 + ht2, ht_gap = unit(2, "cm"))
ignored=dev.off()
include_graphics(role_heatmap_png)
```

## Signaling changes in cells types

```{r}
all_celltypes <- sort(as.character(unique(c(cur_list[[1]]@idents, cur_list[[2]]@idents))))
ct=all_celltypes[1]
ct_content=""
for (ct in all_celltypes){
  ct_filename <- gsub("[ :]", "_", ct)
  g = netAnalysis_signalingChanges_scatter(cur_cellchat, idents.use = ct, color.use=c("grey10", color_use)) + 
    theme(aspect.ratio=1) +
    ggtitle(paste0(ct, "(", case_group, " vs. ", ctrl_group, ")"))
  signaling_changes_png=paste0(cur_comp, ".signaling_changes.", ct_filename, ".png")
  ggsave(signaling_changes_png, g, width=6, height=4, units="in", dpi=300)
  ct_content=paste0(ct_content, "\n### ", ct, "\n\n")
  ct_content=paste0(ct_content, getFigure(signaling_changes_png))

  gdata=g$data
  ct_csv=paste0(cur_comp, ".signaling_changes.", ct_filename, ".csv")
  write.csv(gdata, ct_csv, row.names=TRUE)
  ct_content=paste0(ct_content, getPagedTable(ct_csv, row.names=1))
}
ct_rmd=paste0(cur_comp, ".signaling_changes.rmd")
writeLines(ct_content, ct_rmd)
```

```{r child=ct_rmd}
```

```{r eval=FALSE, include=FALSE}
## Altered signaling based on functional or structural similarity

#This analysis quantifies the similarity between pathways and highligths potential alterations.   

### Signaling groups based on functional similarity

#High degree of functional similarity indicates major senders and receivers are similar, and it can be interpreted as the two signaling pathways or two ligand-receptor pairs exhibit similar and/or redundant roles.
#For each group, the pathways are plotted in a UMAP.   

cur_cellchat <- computeNetSimilarityPairwise(cur_cellchat, type = "functional")
cur_cellchat <- netEmbedding(cur_cellchat, type = "functional")
cur_cellchat <- netClustering(cur_cellchat, type = "functional")

g = netVisual_embeddingPairwise(cur_cellchat, type = "functional", label.size = 3.5) + theme(aspect.ratio=1)
functional_similarity_png=paste0(file_prefix, ".functional_similarity.png")
ggsave(functional_similarity_png, g, width=7, height=6, units="in", dpi=300)
include_graphics(functional_similarity_png)

## Signaling groups based on structure similarity
cur_cellchat <- computeNetSimilarityPairwise(cur_cellchat, type = "structural")
cur_cellchat <- netEmbedding(cur_cellchat, type = "structural")
cur_cellchat <- netClustering(cur_cellchat, type = "structural")

g = netVisual_embeddingPairwise(cur_cellchat, type = "structural", label.size = 3.5) + theme(aspect.ratio=1)
structural_similarity_png=paste0(file_prefix, ".structural_similarity.png")
ggsave(structural_similarity_png, g, width=7, height=6, units="in", dpi=300)
include_graphics(structural_similarity_png)

### Pathway distance

#This plot represents the distance difference of each pathway between the groups from previous UMAP. Larger distance implies larger difference of communication networks between the groups.   
#Only the overlapping pathways are used in this analysis.   

g = rankSimilarity(cur_cellchat, type = "functional", title = "Functional similarity")
similarity_functional_png=paste0(file_prefix, ".functional_distance.png")
ggsave(similarity_functional_png, g, width=6, height=12, units="in", dpi=300)
include_graphics(similarity_functional_png)

g = rankSimilarity(cur_cellchat, type = "structural", title = "Structural similarity")
similarity_structural_png=paste0(file_prefix, ".structural_distance.png")
ggsave(similarity_structural_png, g, width=6, height=12, units="in", dpi=300)
include_graphics(similarity_structural_png)
```

## Identify signaling ligand-receptor pairs

The bubble plot shows the overall signaling changes between these cell types.
The p value refers to the significant communication probabilities, not comparison between the groups.   

```{r}
source_ct=all_celltypes[1]
bubble_content=""
for (source_ct in all_celltypes){
  ct_filename <- gsub("[ :]", "_", source_ct)

  bubble_content=paste0(bubble_content, "\n### ", source_ct, "\n\n")

  g = netVisual_bubble( cur_cellchat, 
                        sources.use = source_ct, 
                        comparison = c(1, 2), 
                        max.dataset = 2, 
                        title.name = paste0("Increased signaling in ", case_group),
                        angle.x = 45, 
                        remove.isolate = T)
  bubble_png=paste0(file_prefix, ".bubble.", ct_filename, ".up_in_", case_group, ".png")
  num_names=length(unique(g$data$interaction_name_2))
  height = min(50, num_names*0.1+5)
  ggsave(bubble_png, g, width=9, height=height, units="in", dpi=300, limitsize = FALSE)

  bubble_content=paste0(bubble_content, "\n#### Increased\n\n")
  bubble_content=paste0(bubble_content, getFigure(bubble_png))

  g = netVisual_bubble( cur_cellchat, 
                        sources.use = source_ct, 
`                       comparison = c(1, 2), 
                        max.dataset = 1, 
                        title.name = paste0("Decreased signaling in ", case_group),
                        angle.x = 45, 
                        remove.isolate = T)
  bubble_png=paste0(file_prefix, ".bubble.", ct_filename, ".down_in_", case_group, ".png")
  num_names=length(unique(g$data$interaction_name_2))
  height = min(50, num_names*0.1+5)
  ggsave(bubble_png, g, width=9, height=height, units="in", dpi=300, limitsize = FALSE)

  bubble_content=paste0(bubble_content, "\n#### Decreased\n\n")
  bubble_content=paste0(bubble_content, getFigure(bubble_png))
}
bubble_rmd=paste0(cur_comp, ".bubble_plot.rmd")
writeLines(bubble_content, bubble_rmd)
```

```{r child=bubble_rmd}
```
