---
title: "Variant Calling QC"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::readthedown:
    toc_depth: 4
    code_folding: hide
    number_sections: true
---

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo=TRUE, 
  include=TRUE, 
  warning=FALSE, 
  message=FALSE, 
  results="asis"
)

library(DT)
library(data.table)
library(ggplot2)
library(vcfR)
library(plotly)

source('reportFunctions.R')
```

```{r}
filelist <- read.table("fileList1.txt", header = F, stringsAsFactors = F)
sample_names <- filelist$V2

vcf_files <- filelist$V1
filter_stats_files <- paste0(filelist$V1, ".filtering.stats")

filterdepth_file <- read.table("fileList2.txt", header = F, stringsAsFactors = F)$V1[1]
filteraf_files <- rev(read.table("fileList3.txt", header = F, stringsAsFactors = F)$V1)

sample_count <- length(sample_names)
```

# Input files

```{r}
input <- data.frame(Sample_Names = sample_names, 
    mutect2_Filter_Stats_Files = filter_stats_files, mutect2_VCF_Files = vcf_files)

print_table(input, byDT=TRUE, row.names=FALSE)
```

```{r}
# Collect data for all samples
all_fdr_stats <- list()
all_filter_stats <- list()
all_filter_step <- list()
all_ti_tv_ratios <- data.frame(Sample = character(), Filter=character(), TI_TV_Ratio = numeric(), stringsAsFactors = FALSE)
all_snp_types <- data.frame(SNP_Type = character(), Count = integer(), Sample = character(), Filter=character(), Percentage = numeric(), stringsAsFactors = FALSE)
all_af_data <- data.frame(Minor_AF = numeric(), Sample = character(), Filter=character(), stringsAsFactors = FALSE)

# Load files once
filterdepth_df <- fread(filterdepth_file, data.table = F, skip="#CHROM", header = TRUE)

filter_af_list <- lapply(filteraf_files, function(x) {
    res = fread(x, data.table = F, comment.char="#", header = TRUE)
    return(res)
})
names(filter_af_list) <- gsub(".+freq", "Exon_MAF<", gsub("\\.filtered\\.tsv$", "", basename(filteraf_files)))

all_variants=c(list("Depth"=filterdepth_df), filter_af_list)

i=1
for (i in 1:length(sample_names)){
    # Global FDR
    fdr_stats <- data.frame(Sample = sample_names[i], FDR = NA)
    stats_lines <- readLines(filter_stats_files[i])
    fdr_line <- stats_lines[grep("fdr=", stats_lines)]
    fdr_stats[1, "FDR"] <- as.numeric(gsub("#<METADATA>fdr=", "", fdr_line))
    all_fdr_stats[[i]] <- fdr_stats
    
    # Filter Statistics (FP, FDR, FN, FNR)
    data_start <- which(!grepl("^(#|\\s*$)", stats_lines))[1]
    f1.1 <- fread(filter_stats_files[i], skip = data_start - 1, header = TRUE, data.table = FALSE)
    f1.1$Sample <- sample_names[i]
    all_filter_stats[[i]] <- f1.1
    
    # vcf 
    vcf <- read.vcfR(vcf_files[i], convertNA = T, verbose = F)
    vcf@gt <- vcf@gt[, 1:2]
    
    # Variants removed by pipeline filters
    filter_step <- data.frame(lapply(all_variants, function(x) sum(x[,sample_names[i]] != "./.")))
    filter_step$Sample <- sample_names[i]
    
    all_filter_step[[i]] <- filter_step
    
    # SNP Types and Ti/Tv ratio
    sname=names(all_variants)[1]
    for(sname in names(all_variants)){
        filter_af <- all_variants[[sname]]
        filter_pass <- filter_af[filter_af[,sample_names[i]] != "./.",]
        if (nrow(filter_pass) == 0) {
            all_ti_tv_ratios <- rbind(all_ti_tv_ratios, data.frame(Sample = sample_names[i], "Filter"=sname, TI_TV_Ratio = NA_real_, stringsAsFactors = FALSE))
            all_af_data <- rbind(all_af_data, data.frame(Minor_AF = NA_real_, Sample = sample_names[i], "Filter"=sname, stringsAsFactors = FALSE))
        } else {
            filter_pass_snps <- paste0(filter_pass[,4], ">", filter_pass[,5])
            filter_pass_snps <- filter_pass_snps[filter_pass[,4] %in% c("A", "C", "G", "T") & filter_pass[,5] %in% c("A", "C", "G", "T")]

            filter_pass_snps_df <- as.data.frame(table(filter_pass_snps)) |>
                dplyr::rename(SNP_Type = 1, Count = 2)
            filter_pass_snps_df$Sample <- sample_names[i]
            filter_pass_snps_df$Filter <- sname
            filter_pass_snps_df$Percentage <- (filter_pass_snps_df$Count / sum(filter_pass_snps_df$Count)) * 100
            all_snp_types <- rbind(all_snp_types, filter_pass_snps_df)

            # Transition/transversion ratio
            ti_count <- sum((table(filter_pass_snps)[c("A>G", "G>A", "C>T", "T>C")]), na.rm = T)
            tv_count <- sum((table(filter_pass_snps)[c("A>C", "C>A", "A>T", "T>A", "C>G", "G>C", "G>T", "T>G")]), na.rm=T)
            ti_tv_ratio <- ti_count / tv_count
            all_ti_tv_ratios <- rbind(all_ti_tv_ratios, data.frame(Sample = sample_names[i], Filter=sname, TI_TV_Ratio = ti_tv_ratio, stringsAsFactors = FALSE))

            # Minor Allele Frequency Distribution
            filter_pass_af <- do.call(rbind,strsplit(filter_pass[,sample_names[i]], ":"))[,3]
            filter_pass_af_minor <- data.frame(Minor_AF = as.numeric(gsub(".*,", "", filter_pass_af)), Filter=sname, Sample = sample_names[i])
            all_af_data <- rbind(all_af_data, filter_pass_af_minor)
        }
    }
}

# Combine all lists into dataframes
all_fdr_stats <- rbindlist(all_fdr_stats, fill=TRUE)
all_filter_stats <- rbindlist(all_filter_stats, fill=TRUE)
all_filter_step <- rbindlist(all_filter_step, fill=TRUE)

filter_level=c("Depth", "Exon_MAF<1.0", "Exon_MAF<0.1", "Exon_MAF<0.01", "Exon_MAF<0.001")
all_ti_tv_ratios$Filter=factor(all_ti_tv_ratios$Filter, levels=filter_level)

sample_x_axis_theme <- if (sample_count > 30) {
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
} else {
    theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
          axis.title.x = element_blank())
}

# Reshape filter_step table: rows are Step, columns are samples with Count
all_filter_step <- all_filter_step |> 
    tibble::column_to_rownames("Sample")
colnames(all_filter_step) <- gsub("MAF.", "MAF<", colnames(all_filter_step))
```

# Mutect2 Global FDR

The FDR is the ratio of expected false positives to the total number of variants passing the filters. A lower FDR indicates higher confidence in the variant calls.

```{r}
print_table(all_fdr_stats, byDT=TRUE, row.names = F)
```

## Mutect2 Global FDR Barplot

```{r}
data_fdr <- all_fdr_stats
data_fdr$Sample <- factor(data_fdr$Sample, levels = data_fdr$Sample[order(data_fdr$FDR, decreasing = TRUE)])
p0 <- ggplot(data_fdr, aes(x = Sample, y = FDR, fill = Sample,
        text = paste0("Sample: ", Sample, "<br>FDR: ", round(FDR, 4)))) +
    geom_col(show.legend = FALSE) +
    labs(y = "FDR") +
    theme_bw() +
    sample_x_axis_theme


ggplotly(p0, tooltip = "text")
```

# Per-sample Beta-binomial and Binomial Clusters

<details><summary>Click to expand</summary>

```{r}
for(i in seq_along(sample_names)){
    cat("## ", sample_names[i], "\n\n")
    stats_lines <- readLines(filter_stats_files[i], warn = FALSE)
    matches <- grep("Background beta-binomial|High-AF beta-binomial|Binomial cluster", stats_lines, value = TRUE)
    if(length(matches) == 0){
        cat("No matches found\n\n")
    } else {
        cat("```\n")
        cat(paste(matches, collapse = "\n"))
        cat("\n```\n\n")
    }
}
```
</details>

# Mutect2 Filter Statistics

- FP: False positives
- FDR: False discovery rate
- FN: False negatives
- FNR: False negative rate

```{r}
print_table(data.frame(all_filter_stats), byDT=TRUE, row.names = F)
```

## Mutect2 Filter Statistics - FP by Filter

```{r}
all_filter_stats_long <- melt(all_filter_stats, id.vars = c("Sample", "filter"), measure.vars = c("FP", "FN"),
    variable.name = "Metric", value.name = "Count")

data_fp <- subset(all_filter_stats_long, Metric == "FP")
data_fp$Sample <- factor(data_fp$Sample, levels = unique(data_fp$Sample[order(data_fp$Count, decreasing = TRUE)]))

p_fp <- ggplot(data_fp, aes(x = Sample, y = Count, fill = Sample,
        text = paste0("Sample: ", Sample, "<br>Filter: ", filter, "<br>FP: ", round(Count, 2)))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~ filter, scales = "free_y", ncol = 3) +
    labs(x = "Sample", y = "FP") +
    theme_minimal() +
    theme(  axis.text.x = element_blank(), 
            axis.ticks.x = element_blank(), 
            strip.text = element_text(size = 10))

ggplotly(p_fp, tooltip = "text")
```

## Mutect2 Filter Statistics - FN by Filter

```{r}
data_fn <- subset(all_filter_stats_long, Metric == "FN")
data_fn$Sample <- factor(data_fn$Sample, levels = unique(data_fn$Sample[order(data_fn$Count, decreasing = TRUE)]))

p_fn <- ggplot(data_fn, aes(x = Sample, y = Count, fill = Sample)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~ filter, scales = "free_y", ncol = 3) +
    labs(x = "Sample", y = "FN") +
    theme_minimal() +
    theme(  axis.text.x = element_blank(), 
            axis.ticks.x = element_blank(), 
            strip.text = element_text(size = 10))

ggplotly(p_fn, tooltip = "text")
```

# Variants Passing Pipeline Filters

## Number of variants passing filter

- Depth: minNormalDepth >= 8, minTumorDepth >= 10, minMinorAlleleDepth >= 3
- Exon_MAF: exonic or splicing variants with MAF < 1.0/0.1/0.01/0.001 in population databases TopMed, ExAC, gnomAD, 1000G

```{r}
print_table(all_filter_step, byDT=TRUE, row.names = T)
```

```{r}
all_filter_step_long <- melt(all_filter_step |> tibble::rownames_to_column("Sample"), 
                            id.vars = "Sample", 
                            variable.name = "Filter", 
                            value.name = "Count")
all_filter_step_long$Filter=factor(all_filter_step_long$Filter, levels=filter_level)                            

filter_names=levels(all_filter_step_long$Filter)

n_samples=length(unique(all_filter_step_long$Sample))
n_col=ceiling(sqrt(n_samples))
n_row=ceiling(n_samples/n_col)

filter_content=""

filter_name = filter_names[1]
for(filter_name in filter_names){
    filter_content = paste0(filter_content, "\n## Filter by: ", filter_name, "\n\n")

    cur_step = all_filter_step_long |> dplyr::filter(Filter == filter_name)
    cur_step$Sample <- factor(cur_step$Sample, levels = cur_step$Sample[order(cur_step$Count, decreasing = TRUE)])

    g=ggplot(cur_step, aes(x=Sample, y=Count)) + 
        geom_bar(stat="identity", fill="steelblue", width=0.5) +
        labs(y = "Number of variants") +
        ggtitle(paste0("Filter: ", filter_name)) +
        theme_bw() + 
        theme(  plot.title = element_text(hjust = 0.5, size=10),
                strip.background = element_blank()) +
        sample_x_axis_theme

    cur_prefix = paste0("filter_", gsub("[<> ]", "_", filter_name))

    cur_png = paste0(cur_prefix, ".png")
    ggsave(cur_png, plot = g, width = 4, height = 3, units = "in", dpi = 300, bg = "white")
    filter_content = paste0(filter_content, getFigure(cur_png))

    cur_snp_types = all_snp_types |> dplyr::filter(Filter == filter_name) |> dplyr::select(-Filter)

    filter_content = paste0(filter_content, "\n### SNP Types - Count\n\n")

    snp_types_count <- dcast(cur_snp_types, SNP_Type ~ Sample, value.var = "Count", fun.aggregate = sum)
    colnames(snp_types_count)[-1] <- paste0(colnames(snp_types_count)[-1], "_Count")
    cur_count_csv = paste0(cur_prefix, ".snp_types_count.csv")
    write.csv(snp_types_count, cur_count_csv)
    filter_content = paste0(filter_content, getTable(cur_count_csv))

    g=ggplot(cur_snp_types, aes(x=SNP_Type, y=Count, fill=SNP_Type)) + 
        geom_bar(stat="identity", position=position_dodge(width=0.5)) +
        facet_wrap(~Sample, ncol=n_col, nrow=n_row) +
        labs(y = "Number of variants") +
        ggtitle(paste0("Filter: ", filter_name)) +
        theme_bw() + 
        theme(  plot.title = element_text(hjust = 0.5, size=10),
                strip.background = element_blank()) +
        sample_x_axis_theme
    cur_count_png = paste0(cur_prefix, ".snp_types_count.png")
    ggsave(cur_count_png, plot = g, width = max(4, n_col * 2) + 2, height = max(4, n_row * 2), units = "in", dpi = 300, bg = "white")
    filter_content = paste0(filter_content, getFigure(cur_count_png))

    filter_content = paste0(filter_content, "\n### SNP Types - Percentage\n\n")

    snp_types_percentage <- dcast(cur_snp_types, SNP_Type ~ Sample, value.var = "Percentage", fun.aggregate = mean)
    colnames(snp_types_percentage)[-1] <- paste0(colnames(snp_types_percentage)[-1], "_Percentage")
    cur_percentage_csv = paste0(cur_prefix, ".percentage.csv")
    write.csv(snp_types_percentage, cur_percentage_csv)
    filter_content = paste0(filter_content, getTable(cur_percentage_csv))

    g=ggplot(cur_snp_types, aes(x=SNP_Type, y=Percentage, fill=SNP_Type)) + 
        geom_bar(stat="identity", position=position_dodge(width=0.5)) +
        facet_wrap(~Sample, ncol=n_col, nrow=n_row) +
        labs(y = "Percentage of variants") +
        ggtitle(paste0("Filter: ", filter_name)) +
        theme_bw() + 
        theme(  plot.title = element_text(hjust = 0.5, size=10),
                strip.background = element_blank()) +
        sample_x_axis_theme
    cur_percentage_png = paste0(cur_prefix, ".snp_types_percentage.png")
    ggsave(cur_percentage_png, plot = g, width = max(4, n_col * 2) + 2, height = max(4, n_row * 2), units = "in", dpi = 300, bg = "white")
    filter_content = paste0(filter_content, getFigure(cur_percentage_png))

    filter_content = paste0(filter_content, "\n### SNP Types - Percentage Stacked Barplot\n\n")

    g <- ggplot(cur_snp_types, aes(x=Sample, y=Percentage, fill=SNP_Type)) +
        geom_col(position = "stack") +
        ggtitle(paste0("Filter: ", filter_name)) +
        scale_y_continuous(labels = function(x) paste0(x, "%")) +
        labs(y = "Percentage of variants", fill = "SNP Type") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5, size=10)) +
        sample_x_axis_theme
    cur_stacked_percentage_png = paste0(cur_prefix, ".snp_types_stacked_percentage.png")
    ggsave(cur_stacked_percentage_png, plot = g, width = max(4, n_samples * 0.5) + 1, height = 4, units = "in", dpi = 300, bg = "white")
    filter_content = paste0(filter_content, getFigure(cur_stacked_percentage_png))
}
filter_rmd="variants_passing_filters.Rmd"
writeLines(filter_content, con = filter_rmd)
```

```{r child=filter_rmd}
```

## Transition/Transversion Ratio

The transition/transversion (Ti/Tv) ratio in human genome is typically 2.

```{r}
all_ti_tv_ratios_wide <- dcast(all_ti_tv_ratios, Sample ~ Filter, value.var = "TI_TV_Ratio", fun.aggregate = mean)
print_table(all_ti_tv_ratios_wide, byDT=TRUE, row.names = F)
```

### Transition/Transversion Ratio - Barplot

```{r}
g <- ggplot(all_ti_tv_ratios, aes(x=Sample, y=TI_TV_Ratio)) +
    geom_bar(width=0.5, stat="identity") +
    facet_wrap(Filter~., ncol=1) +
    labs(y = "Transition/Transversion Ratio") +
    theme_bw() +
    theme(  plot.title = element_blank(),
            strip.background = element_blank(),
            strip.text = element_text(face="bold")) +
    sample_x_axis_theme
cur_ti_tv_png = "ti_tv_ratio.png"
ggsave(cur_ti_tv_png, plot = g, width = max(3, n_samples * 0.5) + 1, height = 8, units = "in", dpi = 300, bg = "white")
include_graphics(cur_ti_tv_png)
```

# Allele fractions of alternate alleles in the tumor

The distribution of allele fractions of alternate alleles among the PASS variants across all samples.

```{r}
af_content=""
for(filter_name in filter_names) {
    af_content = paste0(af_content, "\n## Filter by: ", filter_name, "\n\n")

    cur_prefix = paste0("filter_", gsub("[<> ]", "_", filter_name))

    cur_af_data = all_af_data |> dplyr::filter(Filter == filter_name) |> dplyr::select(-Filter)

    af_density_data <- do.call(rbind, lapply(split(cur_af_data[!is.na(cur_af_data$Minor_AF), ], cur_af_data$Sample[!is.na(cur_af_data$Minor_AF)]), function(df) {
        dens <- density(df$Minor_AF, na.rm = TRUE)
        data.frame(Sample = df$Sample[1], Minor_AF = dens$x, Density = dens$y, stringsAsFactors = FALSE)
    }))
    g <- ggplot(af_density_data, aes(x = Minor_AF, y = Density, color = Sample)) +
        geom_line() +
        ggtitle(paste0("Filter: ", filter_name)) +
        labs(x = "Allele fractions of alternate alleles", y = "Density") +
        theme_bw() + 
        theme(plot.title = element_text(hjust = 0.5),
            strip.background = element_blank(),
            strip.text = element_text(size=10))
    cur_af_plot_png = paste0("af_density_", cur_prefix, ".png")
    ggsave(cur_af_plot_png, plot = g, width = 5, height = 3, units = "in", dpi = 300, bg = "white")

    af_content = paste0(af_content, getFigure(cur_af_plot_png))
}
af_rmd="af_distribution.Rmd"
writeLines(af_content, con = af_rmd)
```

```{r child=af_rmd}
```
