```{r}
load_install<-function(library_name, library_sources=library_name){
  if(!require(library_name, character.only = T)){
    BiocManager::install(library_sources, ask=FALSE)
  }
  library(library_name, character.only = T)
}

load_install("ComplexHeatmap")
load_install("clusterProfiler")
load_install("readr")
load_install("gplots")
load_install("viridis")
load_install("circlize")
load_install("ggplot2")
load_install("ggpubr")

if(!exists("lib_folder")){
  lib_folder="https://raw.githubusercontent.com/shengqh/ngsperl/master/lib"
}
source(paste0(lib_folder, "/CQS/reportFunctions.R"))
source(paste0(lib_folder, "/Report/gene_exp_functions.R"))
```

```{r}
#make sure all required variables are defined
stopifnot(exists("prefix"))

#target_genes is a list of genes that will be the testing set that we will look at where the background_genes is the full list of genes that were tested to get the target genes.
stopifnot(exists("target_genes"))
stopifnot(exists("background_genes"))

#gene_exp_file should be the database to compare to (rows = genes, columns = celltypes/tissues/samples) WILL NOT BE USED IF ALL gene_exp OPTIONS ARE 'F'
#gene_exp_file = "/nobackup/h_cqs/references/tissue_specific/20230627_tabula_logNorm_averageExpression_matrix_by_tissueOnly.txt.gz"
stopifnot(exists("gene_exp_file"))
if(!file.exists(gene_exp_file)) stop(paste0("gene_exp_file does not exist: ", gene_exp_file))

#those variables will be used in ID conversion by clusterProfiler::bitr if fromType != gene_exp_toType
stopifnot(exists("fromType"))
stopifnot(exists("gene_exp_toType"))
stopifnot(exists("OrgDb"))
load_install(OrgDb)

stopifnot(exists("perform_gene_exp_5perc"))

has_background_genes = !all(is.null(background_genes))

if(!exists("draw_gene_exp_heatmap_5perc_global")){
  draw_gene_exp_heatmap_5perc_global=FALSE
}

draw_gene_exp_heatmap_5perc_global = perform_gene_exp_5perc & draw_gene_exp_heatmap_5perc_global & has_background_genes
```

## Gene expression in tissue specific data

Read gene expression data from `r gene_exp_file` and filter out genes that are not expressed in any of the samples.

```{r}
geneExprTab<-read.delim(gene_exp_file, sep = "\t", header = T, as.is = T, stringsAsFactors = F, row.names = 1)
the_sums<-apply(geneExprTab, 1 ,sum)
geneExprTabR<-geneExprTab[which(the_sums>0),]

if(has_background_genes){
  stopifnot(all(target_genes %in% background_genes))

  if(fromType != gene_exp_toType){
    bg = clusterProfiler::bitr(
      background_genes,
      fromType = fromType,
      toType = gene_exp_toType,
      OrgDb = OrgDb
    )
    colnames(bg)<-c("FROM","TO")
    bg<-bg[bg$TO %in% row.names(geneExprTabR),]
    bg<-bg[!duplicated(bg$FROM),]
  }else{
    cur_background_genes<-background_genes[which(background_genes %in% row.names(geneExprTabR))]
    bg=data.frame(FROM=cur_background_genes, TO=cur_background_genes)
  }
}

if(has_background_genes){
  eg=bg[match(target_genes, bg$FROM),] %>% dplyr::filter(!is.na(TO))
}else{
  if(fromType != gene_exp_toType){
    eg = clusterProfiler::bitr(
      target_genes,
      fromType = fromType,
      toType = gene_exp_toType,
      OrgDb = OrgDb
    )
    colnames(eg)<-c("FROM","TO")
    eg<-eg[eg$TO %in% row.names(geneExprTabR),]
    eg<-eg[!duplicated(eg$FROM),]
  }else{
    cur_target_genes<-cur_target_genes[which(cur_target_genes %in% row.names(geneExprTabR))]
    eg=data.frame(FROM=cur_target_genes, TO=cur_target_genes)
  }
}
```

There are `r nrow(eg)` target genes found in gene expression table.

```{r eval=has_background_genes, echo=FALSE}
cat("And there are", nrow(bg), "background genes found in gene expression table.\n")
```

```{r}
heatmap_width=14
heatmap_height=10
```

### Heatmap of targeted genes

Heatmap of gene expression of all of the significant genes in target set. 
Plotted Z-score is calculated for each gene (row) in the heatmap.

```{r echo = T, fig.height=heatmap_width, fig.width=heatmap_height}
tdf<-geneExprTabR[which(row.names(geneExprTabR) %in% eg$TO),]
tdf=calc_z_scores(tdf)

col_fun = colorRamp2(c(min(tdf), 0, max(tdf)), c("blue", "white", "red"))
p=Heatmap(tdf,
          name="z score of logNorm\ngene expression",
          col = col_fun,
          cluster_rows=TRUE,
          cluster_columns=TRUE,
          show_row_names=FALSE)
rm(tdf)

heatmap_png=paste0(prefix, "_geneExpression_heatmap.png")
png(heatmap_png, width = heatmap_width, height = heatmap_height, units = "in", res = 300)
plot(p)
ignored=dev.off()
rm(p)
include_graphics(heatmap_png)
```

```{r, eval=perform_gene_exp_5perc, echo=FALSE}
cat("\n\n### Top 5% expressed genes\n\n")
cat("\n\nFor each tissue, genes with 0 expression were removed, the remaining genes ranked, then the top 5% of genes by rank were assigned to the tissue as a top 5% gene.Top 5% calculation was done only in the genes that are shared between Tabula and the INPUT protein dataset.\n\n")
```

```{r, eval=perform_gene_exp_5perc}
if(has_background_genes){
  top5perc<-top_5perc_genes(geneExprTabR[which(row.names(geneExprTabR) %in% bg$TO),])
}else{
  top5perc<-top_5perc_genes(geneExprTabR)
}

top5perc_genes<-unique(unlist(top5perc))
```

```{r, eval=draw_gene_exp_heatmap_5perc_global, echo=FALSE}
cat("\n\n#### Heatmap of top 5% genes\n\n")
cat("\nPlotted Z-score is calculated for each gene (row) in the heatmap. Row side colors are whether the gene is in the target genes (grey = not in target set, red = in target set).\n\n")
```

```{r, eval=draw_gene_exp_heatmap_5perc_global}
tdf<-geneExprTabR[which(row.names(geneExprTabR) %in% top5perc_genes),]
tdf=calc_z_scores(tdf)

is_target=ifelse(row.names(tdf) %in% eg$TO, "Target", "Background")
col = list(Gene = c( "Target" = "red", "Background" = "darkgrey"))
ha = rowAnnotation(Gene= is_target,col=col)

col_fun = colorRamp2(c(min(tdf), 0, max(tdf)), c("blue", "white", "red"))
p=Heatmap(tdf,
          name="z score of logNorm\ngene expression",
          col = col_fun,
          cluster_rows=TRUE,
          cluster_columns=TRUE,
          show_row_names=FALSE,
          show_row_dend=FALSE,
          right_annotation=ha)
rm(tdf)

global_perc5_heatmap_png=paste0(prefix, "_geneExpression_5perc_heatmap_global.png")
png(global_perc5_heatmap_png, width = heatmap_width, height = heatmap_height, units = "in", res = 300)
plot(p)
ignored=dev.off()
rm(p)
include_graphics(global_perc5_heatmap_png)
```

```{r eval=perform_gene_exp_5perc, echo=FALSE}
cat("\n\n#### Numbers in the top 5% expressed genes among the samples in the gene expression data\n\n")
```

```{r}
ratio_width=12
ratio_height=6
```

```{r eval=perform_gene_exp_5perc, fig.height=ratio_width, fig.width=ratio_height}
cou_rat_tab<-counts_and_ratio_top5perc(top5perc = top5perc, sig_proteins = eg$TO)
cou_rat_tab<-cou_rat_tab[order(cou_rat_tab$Count_sig_proteins_in_top5perc, decreasing = F),]
cou_rat_tab$Tabula_group<-factor(cou_rat_tab$Tabula_group, levels = unique(cou_rat_tab$Tabula_group))
  
write.table(cou_rat_tab, paste0(prefix, "_geneExpression_5perc_counts_ratio.tsv"), sep = "\t", col.names = T, row.names = F, quote = F)
print_table(cou_rat_tab, byDT=TRUE, row.names=FALSE)

## Plot counts
bar_plot_onlygenesExprGT0<- ggplot(cou_rat_tab, aes(x = Tabula_group, y = Count_sig_proteins_in_top5perc)) +
                            geom_bar(stat = "identity", fill = "steelblue")+
                            ggpubr::rotate()+
                            theme_bw()+
                            theme(axis.text.y=element_text(size=7))+ 
                            ggtitle("Number of significant genes \n in top 5% expressed genes")

## Plot percentages
cou_rat_tab<-cou_rat_tab[order(cou_rat_tab$Ratio_sigProteins_in_NumTop5PercGenes, decreasing = F),]
cou_rat_tab$Tabula_group<-factor(cou_rat_tab$Tabula_group, levels = unique(cou_rat_tab$Tabula_group))

bar_plot_onlygenesExprGT0_R<- ggplot(cou_rat_tab, aes(x = Tabula_group, y = Ratio_sigProteins_in_NumTop5PercGenes))+
                              geom_bar(stat = "identity", fill = "salmon")+
                              ggpubr::rotate()+
                              theme_bw()+
                              theme(axis.text.y=element_text(size=7)) + 
                              ggtitle("Ratio of significant genes in top 5% genes divided\n by number of top 5% genes")

g<-ggarrange(bar_plot_onlygenesExprGT0, bar_plot_onlygenesExprGT0_R, nrow = 1, ncol = 2,heights = c(5,5), widths = c(6,6))
top5perc_ratio_png=paste0(prefix, "_geneExpression_5perc_counts_ratio.png")
ggsave(top5perc_ratio_png, g, width = ratio_width, height = ratio_height, units = "in", dpi = 300, bg="white")
include_graphics(top5perc_ratio_png)
cat("\nPlot on the left is the direct counts of overlap between the significant genes and the top 5% expressed genes per tissue, plot on right is the number of direct overlaps divided by the total number of top 5% expressed genes for that tabula tissue.\n")
```

### Tissue specific genes and their distribution across tissues

How many genes reach a z-score of >1 for how many tissues?

```{r}
tdf<-geneExprTabR[which(row.names(geneExprTabR) %in% eg$TO),]
tdf=calc_z_scores(tdf)

genes=data.frame(GENE=eg$FROM, ensemblID=eg$TO)

gene_counts_plot<-proteins_GT1_zscore_diff_tissNum(tdf, genes)

gene_counts_print<-as.data.frame(table(gene_counts_plot$counts))
gene_counts_print$genes<-NA
for(i in 1:nrow(gene_counts_print)){
  gene_counts_print$genes[i]<-paste(gene_counts_plot$Gene_Sym[which(gene_counts_plot$counts==gene_counts_print$Var1[i])], sep = "", collapse = ";")
}
names(gene_counts_print)[1]<-c("Number of Tissues with zscore>1")

print_table(gene_counts_print, byDT=TRUE, row.names=FALSE)
```

What tissues have the most representation in the genes that only have high expression in one tissue (z score > 1 in only one tissue)

```{r}
heatmap_width=12
heatmap_height=10
```

```{r, fig.height=heatmap_width, fig.width=heatmap_height}
ones_df<-gene_counts_plot %>% dplyr::filter(counts == 1)
ones_tdf<-tdf[which(row.names(tdf) %in% ones_df$Ensembl),]
row.names(ones_tdf)<-genes$GENE[match(row.names(ones_tdf),genes$ensemblID)]

col_fun = colorRamp2(c(min(ones_tdf), 0, max(ones_tdf)), c("blue", "white", "red"))
p=Heatmap(ones_tdf,
          name="z score of logNorm\ngene expression",
          col = col_fun,
          cluster_rows=TRUE,
          cluster_columns=TRUE,
          show_row_names=TRUE,
          show_row_dend=FALSE,
          show_column_dend=FALSE,
          column_title = "Gene expression with z-score > 1 in only 1 tissue"
)
rm(ones_df, ones_tdf)

zscore_one_png<-paste0(prefix, "_geneExpression_zscore_one_heatmap.png")
png(zscore_one_png, width = heatmap_width, height = heatmap_height, units = "in", res = 300)
plot(p)
ignored=dev.off()
rm(p)
include_graphics(zscore_one_png)
```
