```{r}
if (exists("reportParametersList") && "tissue_specificity_parameter_list" %in% names(reportParametersList)) {
  for (varName in names(reportParametersList$tissue_specificity_parameter_list)) {
    assign(varName,reportParametersList$tissue_specificity_parameter_list[[varName]])
  }
  
  universeProteins=reportParametersList$background_genes
  selectedProteins=reportParametersList$target_genes
  prefix=reportParametersList$prefix
}

# if (is.data.frame(proteins)) {
#   proteinsTable=proteins
#   proteinsColumn=colnames(proteins)[1]
# } else {
#   proteinsTable=data.frame(proteins=proteins)
#   proteinsColumn="proteins"
# }
```

```{r,echo=FALSE}

#interesed proteins, genesymbol
#selectedProteins

#Optional, selectedProteinsLabels, for heatmap to label protein names
if (!exists("selectedProteinsLabels")) {
  selectedProteinsLabels=NULL
}

#define universeProteins and proteinExpressionCut to do tissue level enrichment analysis
if (!exists("universeProteins")) {
  universeProteins=NULL
}
if (!exists("proteinExpressionCut")) {
  proteinExpressionCut=5
}
if (!exists("tissueLevelEnrichment")) {
  tissueLevelEnrichment=FALSE
}
if (!exists("horizontalHeatmap")) {
  horizontalHeatmap=FALSE
}

#output file name
if (!exists("outputFile")) {
  outputFile=file.path("TissueExpression.Pipeline.Report.html")
}

# #save results as RData
# if (!exists("outputRData")) {
#   outputRDataFile=file.path("TissueExpression.Pipeline.Report.RData")
# }
```

```{r,message=FALSE}
library(TissueEnrich)
library(tidyverse)
library(ComplexHeatmap)
library(circlize)
library(deTS)
```

## Protein Annotation based on Protein Atlas

Data file downloaded from https://www.proteinatlas.org/about/download, "proteinatlas.tsv.zip"

```{r}
#proteinAtlasDatabaseFile="https://www.proteinatlas.org/download/proteinatlas.tsv.zip"
if (!exists("proteinAtlasDatabaseFile")) {
  cat("\nProtein Atlas version 23.0\n\nRelease date: 2023.06.19\n\n")
  proteinAtlasDatabaseFile="/data/cqs/references/protein_atlas/v202306/proteinatlas.tsv"
}
```

```{r}
findProteinInDatabase = function(proteinDatabaseAll, selectedProteins,selectedProteinsLabels=NULL) {
  proteinDatabaseAllOtherNamesList = strsplit(proteinDatabaseAll$`Gene synonym`, ",\ ")
  
  selectedProteinsMatch = data.frame()
  for (i in seq_along(selectedProteins)) {
    selectedProteinOne = selectedProteins[i]
    selectedProteinOneInDbInd = which(proteinDatabaseAll$Gene == selectedProteinOne)
    if (length(selectedProteinOneInDbInd) > 0) {
      #matching "Gene" column
      selectedProteinsMatch = rbind(
        selectedProteinsMatch,
        c(
          selectedProteins[i],
          selectedProteinOneInDbInd[1],
          proteinDatabaseAll[selectedProteinOneInDbInd[1], ][[1]]
        )
      )
    } else {
      #NOT matching "Gene" column
      selectedProteinOneInDbInd = which(
        sapply(proteinDatabaseAllOtherNamesList, function(x)
          selectedProteinOne %in% x)
      )
      if (length(selectedProteinOneInDbInd) > 0) {
        selectedProteinsMatch = rbind(
          selectedProteinsMatch,
          c(
            selectedProteins[i],
            selectedProteinOneInDbInd[1],
            proteinDatabaseAll[selectedProteinOneInDbInd[1], ][[1]]
          )
        )
      } else {
        selectedProteinsMatch = rbind(selectedProteinsMatch, c(selectedProteins[i], NA, NA))
      }
    }
  }
  colnames(selectedProteinsMatch) = c("GeneSymbol", "MatchedInd", "GeneNameInDb")
  if (!is.null(selectedProteinsLabels)) {
    selectedProteinsMatch$GeneLabelForFigure = selectedProteinsLabels
  } else {
    selectedProteinsMatch$GeneLabelForFigure = selectedProteinsMatch$GeneSymbol
  }
  
  proteinDatabase = proteinDatabaseAll[selectedProteinsMatch[, "MatchedInd"], ]
  
  cat("\n\n", paste0(length(selectedProteins)," proteins used in mapping"), "\n\n")
  selectedProteinsNotInGeneColumn = selectedProteinsMatch[which(is.na(selectedProteinsMatch[, 2])), 1]
  if (length(selectedProteinsNotInGeneColumn) > 0) {
    cat("\n", length(selectedProteinsNotInGeneColumn),"proteins can't be mapped to Protein expression database:", paste0(selectedProteinsNotInGeneColumn, collapse = ", "))
  }
  
  return(list(proteinDatabase,selectedProteinsMatch))
}
```

### Mapping to database
```{r,message=FALSE}

proteinDatabaseAll=read_delim(proteinAtlasDatabaseFile)

cat("\nMapping selected proteins to database:\n")
temp=findProteinInDatabase(proteinDatabaseAll, selectedProteins,selectedProteinsLabels)
proteinDatabaseSelectedProteins=temp[[1]]
colnames(proteinDatabaseSelectedProteins)=make.names(colnames(proteinDatabaseSelectedProteins))
selectedProteinsMatch=temp[[2]]

if (!is.null(universeProteins)) {
  #universeProteins defined.
  cat("\nMapping universe proteins to database:\n")

  temp = findProteinInDatabase(proteinDatabaseAll, universeProteins)
  proteinDatabaseUniverse = temp[[1]]
  colnames(proteinDatabaseUniverse)=make.names(colnames(proteinDatabaseUniverse))
  universeProteinsMatch = temp[[2]]
}
```

## Expression Specificity Type

https://www.proteinatlas.org/humanproteome/tissue/tissue+specific

**Defining Tissue-specific Genes**
Tissue-specific genes are defined using the algorithm from the HPA (Uhlén et al. 2015), and can be grouped as follows:

**Tissue Enriched**: At least four-fold higher mRNA level in a particular tissue compared to any other tissue.

**Group Enriched**: At least four-fold higher average mRNA level in a group of 2-5 tissues compared to any other tissue.

**Tissue Enhanced**: At least four-fold higher mRNA level in a particular tissue compared to the average level in all other tissues.

```{r}
get_category<-function(proteins){
  result=rep("Intracellular",nrow(proteins))
  result[grep("membrane proteins",proteins[["Protein.class"]])]="Membrane"
  result[grep("secreted proteins",proteins[["Protein.class"]])]="Secreted"
  names(result)=proteins$Gene
  return(result)
}

#generate secreted proteins type
if (exists("proteinDatabaseUniverse")) {
  proteinDatabaseUniverse$Secreted.Category=get_category(proteinDatabaseUniverse)
}

selectedProteinsCategory=get_category(proteinDatabaseSelectedProteins)
proteinDatabaseSelectedProteins$Secreted.Category=selectedProteinsCategory
```

```{r}
library(RColorBrewer)
#source functions from github
source("https://raw.githubusercontent.com/shengqh/ngsperl/master/lib/CQS/countTableVisFunctions.R")
```

```{r}
nSelected=nrow(proteinDatabaseSelectedProteins)

if (exists("proteinDatabaseUniverse")) {
  nUniverse=nrow(proteinDatabaseUniverse)
}

summarizeCategoricalDataFromProteinDatabase = function(maxCategory = 10,
                                    columnName = "RNA.tissue.specificity") {
  #use variable name as column name in group_by
  dataSelected = proteinDatabaseSelectedProteins %>% group_by(!!sym(columnName)) %>% summarise(n =
                                                                                                 n()) %>% mutate(Category = "Selected") %>% arrange(desc(n))
  dataSelected[which(is.na(dataSelected[, 1])), 1] = "Not Available"
  
  #keep top 10 rows (by n) in dataSelected
  if (nrow(dataSelected) > (maxCategory + 1)) {
    dataSelectedSub = rbind(dataSelected[1:maxCategory, ], c(
      "Others",
      n = sum(dataSelected[(maxCategory + 1):nrow(dataSelected), "n"]),
      Category = "Selected"
    ))
  } else {
    dataSelectedSub = dataSelected
  }
  dataForPlot = dataSelectedSub
  dataForPlot$n = as.numeric(dataForPlot$n)

  if (exists("proteinDatabaseUniverse")) {
    dataUniverse = proteinDatabaseUniverse %>% group_by(!!sym(columnName)) %>% summarise(n =
                                                                                           n()) %>% mutate(Category = "Background")
    dataUniverse[which(is.na(dataUniverse[, 1])), 1] = "Not Available"
    
    temp=(setdiff(dataUniverse[[1]],dataSelectedSub[[1]]))
    if (length(temp)>0) {
          dataUniverseSub = rbind(dataUniverse[which(dataUniverse[[1]] %in% dataSelectedSub[[1]]), ],
                            c(
                              "Others",
                              n = sum(dataUniverse[-which(dataUniverse[[1]] %in% dataSelectedSub[[1]]), "n"]),
                              Category = "Background"
                            ))
    } else {
          dataUniverseSub = dataUniverse
    }

    dataForPlot = rbind(dataSelectedSub, dataUniverseSub)
    dataForPlot$n = as.numeric(dataForPlot$n)
    
    dataForAnalysis = dataSelected %>% left_join(dataUniverse, by = columnName) %>% mutate(Enrichment =
                                                                                             n.x / n.y) %>% arrange(desc(Enrichment))
    
    #run a Hypergeometric test by phyper
    dataForAnalysis = dataForAnalysis %>% mutate(
      pvalue = phyper(n.x - 1, nSelected, nUniverse - nSelected, n.y, lower.tail =
                        FALSE),
      pAdj = p.adjust(pvalue, method = "BH"),
      pAdjScore = -log10(pAdj)
    )
    
    dataForAnalysis=dataForAnalysis %>% slice_max(n=maxCategory,order_by=pAdjScore) %>% mutate(Levels=factor(!!sym(columnName),levels=unique(!!sym(columnName))[order(pAdjScore)]))
    return(list(dataForPlot, dataForAnalysis))
  } else {
    return(list(dataForPlot))
  }
}
```

```{r,results='asis',fig.width=9}
allColumnNames = c(
  "Secreted.Category",
  "RNA.tissue.specificity",
  "RNA.single.cell.type.specificity",
  "RNA.cancer.specificity",
  "RNA.blood.cell.specificity",
  "RNA.blood.lineage.specificity",
  "RNA.cell.line.specificity"
)
maxCategory=10

for (columnName in allColumnNames) {
  cat("\n\n")
  cat(paste0("### ", columnName, "\n\n"))
  cat("\n\n")
  
  temp = summarizeCategoricalDataFromProteinDatabase(maxCategory = maxCategory,
                                                     columnName = columnName)
  dataForPlot = temp[[1]]
  
  p=myggpie(
    data.frame(dataForPlot),
    y = "n",
    fill = columnName,
    facet = "Category",
    transformTable = FALSE
  )
  plot(p)
  
  if (exists("proteinDatabaseUniverse")) {
    dataForAnalysis = temp[[2]]
    p=ggplot(dataForAnalysis, aes(x = pAdjScore, y = Levels)) + geom_col()
    plot(p+theme_bw()+ geom_vline(xintercept = -log10(0.05),linetype="dashed"))
  }
}
```

## Expression Cluster

### Tissue.expression.cluster

```{r,fig.width=9}

maxCategory=10
columnName="Tissue.expression.cluster"

temp=summarizeCategoricalDataFromProteinDatabase(maxCategory = maxCategory,
                                    columnName = columnName)
dataForPlot=temp[[1]]
myggpie(data.frame(dataForPlot),y="n",fill=columnName,facet="Category",transformTable=FALSE)

if (exists("proteinDatabaseUniverse")) {
  dataForAnalysis=temp[[2]]
  p=ggplot(dataForAnalysis,aes(x=pAdjScore,y=Levels)) + geom_col()
  plot(p+theme_bw()+ geom_vline(xintercept = -log10(0.05),linetype="dashed"))
}

```

### Single.cell.expression.cluster
```{r,fig.width=9}

maxCategory=10
columnName="Single.cell.expression.cluster"

temp=summarizeCategoricalDataFromProteinDatabase(maxCategory = maxCategory,
                                    columnName = columnName)
dataForPlot=temp[[1]]
myggpie(data.frame(dataForPlot),y="n",fill=columnName,facet="Category",transformTable=FALSE)

if (exists("proteinDatabaseUniverse")) {
  dataForAnalysis=temp[[2]]
  p=ggplot(dataForAnalysis,aes(x=pAdjScore,y=Levels)) + geom_col()
  plot(p+theme_bw()+ geom_vline(xintercept = -log10(0.05),linetype="dashed"))
}
```

## Expression Pattern and Enrichment in Tissues/Cells
```{r}
nSelected=nrow(proteinDatabaseSelectedProteins)

if (exists("proteinDatabaseUniverse")) {
  nUniverse=nrow(proteinDatabaseUniverse)
}

summarizenTPMDataFromProteinDatabase = function(maxCategory = 20,
                                                columnName = "RNA.tissue.specific.nTPM") {
  columnTypeName = gsub("specific.nTPM", "specificity", columnName)
  
  #split data into multiple rows
  TissueEnrichSelectedProteins = proteinDatabaseSelectedProteins[, c("Gene", columnTypeName, columnName)] %>% separate_rows(!!sym(columnName),
                                                                                                                            sep = ";") %>% separate(!!sym(columnName),
                                                                                                                                                    into = c("Tissue", "nTPM"),
                                                                                                                                                    sep = ": ")
  TissueEnrichSelectedProteins = na.omit(TissueEnrichSelectedProteins)
  
  print(knitr::kable(unique(TissueEnrichSelectedProteins[,c("Gene",columnTypeName)]) %>% group_by(!!sym(columnTypeName)) %>% summarise(Count=n()),caption = "Number of Proteins with Enriched/Enhanced expression"))

  tissueOrder = names(sort(table(TissueEnrichSelectedProteins$Tissue), decreasing = TRUE))
  
  TissueEnrichSelectedProteinsTypeMatrix = pivot_wider(
    TissueEnrichSelectedProteins[, -4],
    names_from = Tissue,
    values_from = !!sym(columnTypeName)
  )
  TissueEnrichSelectedProteinsTypeMatrix = TissueEnrichSelectedProteinsTypeMatrix[, tissueOrder]
  #order row of dataForPlot by values from first column to last column
  TissueEnrichSelectedProteinsTypeMatrix = TissueEnrichSelectedProteinsTypeMatrix[do.call(order, TissueEnrichSelectedProteinsTypeMatrix),]
  
  TissueEnrichSelectedProteinsnTPMMatrix = pivot_wider(
    TissueEnrichSelectedProteins[, -2],
    names_from = Tissue,
    values_from = nTPM,
    values_fn = as.numeric
  )
  TissueEnrichSelectedProteinsnTPMMatrix = TissueEnrichSelectedProteinsnTPMMatrix[, tissueOrder]
  #order row of dataForPlot by values from first column to last column
  TissueEnrichSelectedProteinsnTPMMatrix = TissueEnrichSelectedProteinsnTPMMatrix[do.call(order,-TissueEnrichSelectedProteinsnTPMMatrix),]
  
  #ComplexHeatmap::Heatmap(scale(log(TissueEnrichSelectedProteinsnTPMMatrix)),cluster_rows = FALSE,cluster_columns = FALSE)
  
  if (exists("proteinDatabaseUniverse")) {
    #split data into multiple rows
    TissueEnrichUniverse = proteinDatabaseUniverse[, c("Gene", columnTypeName, columnName)] %>% 
      separate_rows(!!sym(columnName), sep = ";") %>% 
      separate(!!sym(columnName), into = c("Tissue", "nTPM"), sep = ": ")

    TissueEnrichUniverse = na.omit(TissueEnrichUniverse)
    TissueEnrichUniverseTypeMatrix = pivot_wider(
      TissueEnrichUniverse[, -4],
      names_from = Tissue,
      values_from = !!sym(columnTypeName)
    )
    
    temp1 = apply(TissueEnrichSelectedProteinsTypeMatrix, 2, function(x)
      length(which(!is.na(x))))
    temp2 = apply(TissueEnrichUniverseTypeMatrix[, colnames(TissueEnrichSelectedProteinsTypeMatrix)], 2, function(x)
      length(which(!is.na(x))))
    
    dataForAnalysis = data.frame(n.x = temp1, n.y = temp2) %>% 
      rownames_to_column(var = columnName) %>%
      mutate(Enrichment = n.x / n.y) %>% 
      arrange(desc(Enrichment)) %>% 
      mutate(
        pvalue = phyper(n.x - 1, nSelected, nUniverse - nSelected, n.y, lower.tail = FALSE),
        pAdj = p.adjust(pvalue, method = "BH"),
        pAdjScore = -log10(pAdj)
      )

    dataForAnalysis = dataForAnalysis %>% 
      slice_max(n = maxCategory, order_by = pAdjScore) %>% 
      mutate(Levels = factor(!!sym(columnName), levels = unique(!!sym(columnName))[order(pAdjScore)]))

    return(
      list(
        TissueEnrichSelectedProteinsTypeMatrix,
        TissueEnrichSelectedProteinsnTPMMatrix,
        dataForAnalysis
      )
    )
  } else {
    return(
      list(
        TissueEnrichSelectedProteinsTypeMatrix,
        TissueEnrichSelectedProteinsnTPMMatrix
      )
    )
  }
}
  
```


```{r,results='asis',fig.width=9,warning=FALSE}
allColumnNames = c(
  "RNA.tissue.specific.nTPM",
  "RNA.single.cell.type.specific.nTPM",
  "RNA.blood.cell.specific.nTPM",
  "RNA.blood.lineage.specific.nTPM",
  "RNA.cell.line.specific.nTPM"
)
maxCategory=20

for (columnName in allColumnNames) {
  cat("\n\n")
  cat(paste0("### ", gsub(".nTPM","",columnName), "\n\n"))
  cat("\n\n")
  
  result = summarizenTPMDataFromProteinDatabase(maxCategory = maxCategory,
                                                     columnName = columnName)
  
  dataForPlot = result[[1]]
  temp=table(unlist(dataForPlot ))
  col_fun=RColorBrewer::brewer.pal("Set1",n=length(temp))
  names(col_fun)=names(temp)
  p=ComplexHeatmap::Heatmap(dataForPlot,heatmap_legend_param = list(title="Type"),col=col_fun)
  plot(p)
  
  dataForPlot = result[[2]]
  #p=ComplexHeatmap::Heatmap(scale(log(dataForPlot)),cluster_rows = FALSE,cluster_columns = FALSE,heatmap_legend_param = list(title=paste0("scaled ",columnName)))
  dataForPlot[is.na(dataForPlot)]=0
  temp=log2(dataForPlot+1)
  col_fun = colorRamp2(c(0,max(temp)), c("white", "red"))
  p=ComplexHeatmap::Heatmap((temp),cluster_rows = FALSE,cluster_columns = FALSE,heatmap_legend_param = list(title=paste0("Transformed ",columnName)),col=col_fun)
  plot(p)
  
  if (exists("proteinDatabaseUniverse")) {
    dataForAnalysis = result[[3]]
    p=ggplot(dataForAnalysis, aes(x = pAdjScore, y = Levels)) + geom_col()
    plot(p+theme_bw()+ geom_vline(xintercept = -log10(0.05),linetype="dashed"))
  }
}
```

## Expression of selected proteins in tissues from Protein Atlas by TissueEnrich

https://pubmed.ncbi.nlm.nih.gov/30346488/

```{r,fig.width=12}
#| fig.height=max(length(selectedProteins)/10,7)

# if (!exists("selectedTissues")) {
#   selectedTissues=c("Adipose.Tissue","Bone.Marrow","Heart.Muscle","Kidney","Liver","Lung" ,"Lymph.Node","Skeletal.Muscle","Smooth.Muscle","Spleen")
# }


#expression pattern of these genes in the database
e <- new.env()
load(file = system.file("extdata", "combine-expression.rda",package = "TissueEnrich"),envir = e)
#head(dataset[["Protein-Atlas"]]$expressionData)
#dataset$humanGeneMapping
teExpressionData=e$dataset[["Protein-Atlas"]]$expressionData
teExpressionData$Gene <- row.names(teExpressionData)
geneMapping <- e$dataset[["humanGeneMapping"]]

#selectedGenesEnsg=geneMapping[which(geneMapping$Gene.name %in% selectedProteins),]
selectedGenesEnsg=geneMapping[which(geneMapping$Gene.name %in% selectedProteinsMatch$GeneNameInDb),] #match to gene name in Db to match to expression database

teExpressionDataSelectedProteins <- dplyr::left_join(selectedGenesEnsg,teExpressionData,by = "Gene")

teExpressionDataSelectedProteins=as.data.frame(teExpressionDataSelectedProteins[,-1] %>% group_by(Gene.name) %>% summarise_all(mean,na.rm=TRUE))
row.names(teExpressionDataSelectedProteins)=teExpressionDataSelectedProteins$Gene.name
if (exists("selectedTissues")) {
  teExpressionDataSelectedProteins=teExpressionDataSelectedProteins[,selectedTissues]
  heatmapWidth=4
} else {
  teExpressionDataSelectedProteins=teExpressionDataSelectedProteins[,-1] #remove first column, gene names
  heatmapWidth=12
}

dataForPlot=teExpressionDataSelectedProteins
col_fun = colorRamp2(c(0,max(log2(dataForPlot+1))), c("white", "red"))
annotationCol=RColorBrewer::brewer.pal("Set1",n=3)

if (exists("horizontalHeatmap") & horizontalHeatmap) { #make horizontal heatmap and NOT show trees
  ha = columnAnnotation(Secretome= selectedProteinsCategory[row.names(dataForPlot)],
                   col = list(Secretome = c("Secreted" = annotationCol[1],
                                            "Membrane" = annotationCol[2],
                                          "Intracellular" = annotationCol[3])))
} else {
  ha = rowAnnotation(Secretome= selectedProteinsCategory[row.names(dataForPlot)],
                   col = list(Secretome = c("Secreted" = annotationCol[1],
                                            "Membrane" = annotationCol[2], 
                                            "Intracellular" = annotationCol[3])))
}

geneNamesInFigure=data.frame(GeneNameInDb =row.names(dataForPlot)) %>% left_join(selectedProteinsMatch,by="GeneNameInDb") %>%pull("GeneLabelForFigure")

if (exists("horizontalHeatmap") & horizontalHeatmap) { #make horizontal heatmap and NOT show trees
tissueHeatmapResult=Heatmap(t(as.matrix(log2(dataForPlot+1))),name="Log2(TPM+1)",
                              col = col_fun,column_names_rot = 45,
                            cluster_rows=TRUE,cluster_columns =TRUE,
                              show_column_dend =FALSE,show_row_dend =FALSE,
                              width = unit(20, "cm"),
                              height = unit(4, "cm"),
                            top_annotation = ha,
                            column_labels =geneNamesInFigure)
} else {
    tissueHeatmapResult=Heatmap(as.matrix(log2(dataForPlot+1)),name="Log2(TPM+1)",
                             col = col_fun,cluster_rows=TRUE,cluster_columns =TRUE,
                              width = unit(heatmapWidth, "cm"),right_annotation = ha,
                            row_labels =geneNamesInFigure)
}

plot(tissueHeatmapResult)
```

```{r,fig.width=9,eval=FALSE}

# ## Tissue-specific gene enrichment by TissueEnrich: NOT using this

# https://bioconductor.org/packages/release/bioc/vignettes/TissueEnrich/inst/doc/TissueEnrich.html

# ### Enrichment in tissue level

gs<-GeneSet(geneIds=selectedProteinsMatch$GeneSymbol,organism="Homo Sapiens",geneIdType=SymbolIdentifier())

if (exists("universeProteins") & !is.null(universeProteins)) {
  backgroundGenes=GeneSet(geneIds=universeProteinsMatch$GeneSymbol,organism="Homo Sapiens",geneIdType=SymbolIdentifier())
  TissueEnrichOutput<-teEnrichment(inputGenes = gs,backgroundGenes=backgroundGenes)
} else {
  TissueEnrichOutput<-teEnrichment(inputGenes = gs)
}

seEnrichmentOutput<-TissueEnrichOutput[[1]]
enrichmentOutput<-setNames(data.frame(assay(seEnrichmentOutput),row.names = rowData(seEnrichmentOutput)[,1]), colData(seEnrichmentOutput)[,1])
enrichmentOutput$Tissue<-row.names(enrichmentOutput)
#head(enrichmentOutput)

p=ggplot(enrichmentOutput,aes(x=reorder(Tissue,-Log10PValue),y=Log10PValue,label = Tissue.Specific.Genes,fill = Tissue))+
      geom_bar(stat = 'identity')+
      labs(x='', y = '-LOG10(P-Adjusted)')+
      theme_bw()+
      theme(legend.position="none")+
      theme(plot.title = element_text(hjust = 0.5,size = 20),axis.title = element_text(size=15))+
      theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),panel.grid.major= element_blank(),panel.grid.minor = element_blank())
plot(p)
```

```{r,fig.width=9,warning=FALSE,results='asis',eval=FALSE}
# ### Heatmap for Enriched proteins

TissueEnrichProteinsAll=NULL
for (i in seq_along(TissueEnrichOutput[[3]])) {
  temp=data.frame(assay(TissueEnrichOutput[[3]][[i]]),Tissue=names(TissueEnrichOutput[[3]])[i])
  TissueEnrichProteinsAll=rbind(TissueEnrichProteinsAll,temp)
}

# #heatmap by value
# #dataForPlot=pivot_wider(TissueEnrichProteinsAll,names_from = Tissue,values_from = Group)
# dataForPlot=pivot_wider(TissueEnrichProteinsAll,names_from = Tissue,values_from = Group,
#                         values_fn = function(x) case_when(x=="Tissue-Enhanced"~1,
#                                               x=="Group-Enriched"~2,
#                                               x=="Tissue-Enriched"~3,
#                                               TRUE~0))
# dataForPlot[is.na(dataForPlot)]=0
# ComplexHeatmap::Heatmap(dataForPlot[,-1])

dataForPlot=pivot_wider(TissueEnrichProteinsAll,names_from = Tissue,values_from = Group)
tissueOrder=names(sort(table(TissueEnrichProteinsAll$Tissue),decreasing = TRUE))
dataForPlot=dataForPlot[,tissueOrder]
# ComplexHeatmap::Heatmap(dataForPlot)

#order row of dataForPlot by values from first column to last column
dataForPlot=dataForPlot[ do.call(order, dataForPlot), ]
#ComplexHeatmap::Heatmap(dataForPlot)

knitr::kable(unique(TissueEnrichProteinsAll[,c("Gene","Group")]) %>% group_by(Group) %>% summarise(Count=n()),caption = "Number of Proteins with Enriched/Enhanced tissue expression")

#define colors of categical data in ComplexHeatmap
col_fun = structure( RColorBrewer::brewer.pal(3,"Set1")  ,names=c("Tissue-Enriched","Tissue-Enhanced", "Group-Enriched" ))
ComplexHeatmap::Heatmap(dataForPlot,col=col_fun,name = "Tissue Enrichment Type")
```

```{r,results='asis',eval=FALSE}

## Tissue specific Function Enrichment Analysis: NOT using this

#In this analysis, we will do Function Enrichment Analysis in each tissue separately, only using proteins with TPM>=proteinExpressionCut in that tissue (based on Protein Atlas database). At last we summarized the top 3 most significant pathways in each tissue together in one heatmap.


if (!is.null(universeProteins) & exists("tissueLevelEnrichment") & tissueLevelEnrichment) {
  #universeProteins defined. need to do tissue level enrichment analysis
  cat("\n\n")
  cat(
    "## Function enrichment analysis of selected proteins based on expression pattern in tissues from Protein Atlas
"
  )
  cat("\n\n")
  
  temp = findProteinInDatabase(proteinDatabaseAll, universeProteins)
  proteinDatabaseUniverse = temp[[1]]
  universeProteinsMatch = temp[[2]]
  
  proteinDatabaseUniverseEnsg = geneMapping[which(geneMapping$Gene.name %in% universeProteinsMatch$GeneNameInDb), ] #match to gene name in Db to match to expression database
  
  teExpressionDataUniverse <-
    dplyr::left_join(proteinDatabaseUniverseEnsg, teExpressionData, by = "Gene")
  
  teExpressionDataUniverseFormated = as.data.frame(
    teExpressionDataUniverse[, -1] %>% group_by(Gene.name) %>% summarise_all(mean, na.rm =
                                                                               TRUE)
  )
  row.names(teExpressionDataUniverseFormated) = teExpressionDataUniverseFormated$Gene.name
  if (exists("selectedTissues")) {
    teExpressionDataUniverseFormated = teExpressionDataUniverseFormated[, selectedTissues]
  } else {
    teExpressionDataUniverseFormated = teExpressionDataUniverseFormated[, -1] #remove first column, gene names
  }
  
  #tissue level encirhment analysis
  if (ncol(teExpressionDataUniverseFormated) >= 1) {
    warning(paste0(
      "Too many tissues. Only keeping first 10 tissues for enrichment analysis"
    ))
    teExpressionDataUniverseFormated = teExpressionDataUniverseFormated[, 1:10]
  }
  
  
  enrichmentResultTissue = list()
  for (i in 1:ncol(teExpressionDataUniverseFormated)) {
    selectedProteinsTissueOne = row.names(teExpressionDataSelectedProteins)[which(teExpressionDataSelectedProteins[, i] >=
                                                                                    proteinExpressionCut)]
    proteinUniverseInTissueOne = row.names(teExpressionDataUniverseFormated)[which(teExpressionDataUniverseFormated[, i] >=
                                                                                     proteinExpressionCut)]
    enrichmentResultTissue[[i]] = enrichmentByClusterProfiler(
      selectedProteinsTissueOne,
      universe = proteinUniverseInTissueOne,
      modules = c("Reactome", "KEGG", "GO.BP"),
      organism = "hsa",
      fromType = "SYMBOL",
      toType = "ENTREZID",
      OrgDb = "org.Hs.eg.db",
      minGSSize = 3,
      maxGSSize = 1000,
      pvalueCutoff = 1,
      qvalueCutoff = 1,
      showConvertedGeneId = TRUE
    )
    
  }
  names(enrichmentResultTissue)=colnames(teExpressionDataUniverseFormated)
  
  
  #extract enrichment result for heatmap vis
  for (module in c("Reactome", "KEGG", "GO.BP")) {
    dataForPlot=lapply(enrichmentResultTissue,function(x) x[[module]])
    p=makeHeatmapClusterProfilerEnrichment(dataForPlot)
    
  cat("\n\n")
  cat(paste0(
    "### ",module," \n\n"
  ))
  cat("\n\n")
    plot(p)
  }
}
```

## Expression of selected proteins in tissues from GTEx by deTS

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6761978/

```{r,fig.height=6,message=FALSE,results='hide'}

data(GTEx_t_score)
#data(ENCODE_z_score)

tsea_t = tsea.analysis(selectedProteins, GTEx_t_score, ratio = 0.05, p.adjust.method = "bonferroni")
tsea_t=tsea_t[order(tsea_t[,1]),,drop=FALSE]

try(tsea.plot(tsea_t, threshold = 0.05))
tsea_t_summary = tsea.summary(tsea_t)
```

```{r,fig.width=9,message=FALSE}
#| fig.height=max(length(selectedProteins)/10,7)
dataForPlot=scale(GTEx_t_score[intersect(row.names(GTEx_t_score),selectedProteins),])
if (exists("selectedTissuesFordeTS")) {
  dataForPlot=dataForPlot[,selectedTissuesFordeTS]
}
#remove all NA rows generated by scale
dataForPlot=dataForPlot[which(apply(dataForPlot,1,function(x) !all(is.na(x)))),]

col_fun = colorRamp2(c(-max(abs(dataForPlot),na.rm=TRUE),0,max(abs(dataForPlot),na.rm=TRUE)), c("blue","white", "red"))
# annotationCol=RColorBrewer::brewer.pal("Set1",n=3)
haBydeTS = rowAnnotation(Secretome= selectedProteinsCategory[row.names(dataForPlot)],
                   col = list(Secretome = c("Secreted" = annotationCol[1],
                                            "Membrane" = annotationCol[2],
                                            "Intracellular" = annotationCol[3])))
tissueHeatmapResultBydeTS=Heatmap(dataForPlot,name="scaled T stastics",
                             col = col_fun,cluster_rows=TRUE,cluster_columns =TRUE,
                             right_annotation = haBydeTS)

plot(tissueHeatmapResultBydeTS)
```

## Save results

```{r}
# save.image(outputRDataFile)

saveRDS(selectedProteinsCategory,paste0(outputFile,".selectedProteinsCategory.rds"))
saveRDS(teExpressionDataSelectedProteins,paste0(outputFile,".teExpressionDataSelectedProteins.rds"))

```
