---
title: "methylKit tiling-window DMR analysis"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output:
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: hide
    number_sections: true
---

```{css, echo=FALSE}
#content {
  max-width: 1600px;
}
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      results = "asis", fig.width = 7, fig.height = 5)
library(DT)
library(ggplot2)
library(knitr)
library(data.table)

find_one <- function(pattern) {
  matches <- list.files(pattern = pattern, full.names = TRUE)
  if (length(matches) != 1) stop("Expected one file matching ", pattern, "; found ", length(matches))
  matches[[1]]
}

parameters <- fread(find_one("[.]methylkit[.]dmr[.]parameters[.]tsv$"), data.table = FALSE)
samples <- fread(find_one("[.]methylkit[.]dmr[.]samples[.]tsv$"), data.table = FALSE)
dmrs <- fread(find_one("[.]methylkit[.]dmrs[.]tsv$"), data.table = FALSE)
annotated <- fread(find_one("[.]methylkit[.]dmr[.]annotated[.]tsv$"), data.table = FALSE)
annotation_summary <- fread(find_one("[.]methylkit[.]dmr[.]annotation_summary[.]tsv$"), data.table = FALSE)
chromosome_summary <- fread(find_one("[.]methylkit[.]dmr[.]chromosome_summary[.]tsv$"), data.table = FALSE)
parameter_values <- setNames(as.character(parameters$value), parameters$parameter)
comparison <- parameter_values[["comparison"]]
```

# Comparison

```{r}
cat("## ", comparison, "\n\n", sep = "")
```

The methylation calls were summarized in genomic tiling windows before differential testing. Positive methylation differences indicate higher methylation in the treatment group; negative values indicate higher methylation in the control group.

```{r}
overview <- data.frame(
  Metric = c("Control group", "Treatment group", "Control samples", "Treatment samples",
             "Tested regions", "Significant DMRs", "Hyper in treatment", "Hyper in control"),
  Value = c(
    parameter_values[["control_group"]],
    parameter_values[["treatment_group"]],
    sum(samples$role == "control"),
    sum(samples$role == "treatment"),
    parameter_values[["tested_regions"]],
    parameter_values[["significant_dmrs"]],
    if (nrow(dmrs) == 0) 0 else sum(dmrs$meth.diff > 0),
    if (nrow(dmrs) == 0) 0 else sum(dmrs$meth.diff < 0)
  ),
  stringsAsFactors = FALSE
)
kable(overview, row.names = FALSE)
```

## Input samples

```{r}
DT::datatable(samples, rownames = FALSE, options = list(pageLength = 10, scrollX = TRUE))
```

# Differentially methylated regions

```{r}
if (nrow(annotated) == 0) {
  cat("No regions passed the configured methylation-difference and significance thresholds.\n")
} else {
  DT::datatable(annotated, rownames = FALSE, filter = "top",
                options = list(pageLength = 25, scrollX = TRUE))
}
```

## Chromosome summary

```{r}
if (nrow(chromosome_summary) == 0) {
  cat("No significant DMRs were available for a chromosome summary.\n")
} else {
  graph <- ggplot(chromosome_summary,
                  aes(x = reorder(chromosome, count), y = count, fill = direction)) +
    geom_col(position = "stack") +
    coord_flip() +
    labs(x = "Chromosome", y = "DMRs", fill = "Direction") +
    theme_bw()
  print(graph)
}
```

# Genomic annotation

The gene-context classification uses promoter, exon, intron, and intergenic precedence. CpG-island annotation classifies regions as CpG island, shore, or other; islands take precedence over shores when a window overlaps both.

```{r, out.width='49%', fig.show='hold'}
knitr::include_graphics(c(
  find_one("[.]methylkit[.]dmr[.]annotation[.]png$"),
  find_one("[.]methylkit[.]dmr[.]cpg_context[.]png$")
))
```

```{r}
kable(annotation_summary, digits = 2, row.names = FALSE)
```

# Analysis settings

```{r}
kable(parameters, row.names = FALSE)
```
