---
title: "RNA Sequencing Data Analysis Report"
date: "`r format(Sys.time())`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
---

<style type="text/css">
.main-container {
  max-width: 90%;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, child="Functions.Rmd"} 
```
```{r, child="Pipeline.Rmd"} 
```
  
```{r setup, include=FALSE}
library(kableExtra)
library(tidyverse)

files<-read.table("fileList1.txt", header=FALSE, as.is=TRUE)
rownames(files)<-files$V2

rnaseqOptions<-read.table("fileList2.txt", header=FALSE, as.is=TRUE)
rownames(rnaseqOptions)<-rnaseqOptions$V2

taskName<-rnaseqOptions["task_name", 1]
out.width<-rnaseqOptions["out.width", 1]

knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 120),
  message = FALSE,
  warning = FALSE,
  fig.width=8, 
  fig.height=6, 
  out.width=out.width
)


hasFastqc<-any(grepl("fastqc_", rownames(files)))
hasFastqcPosttrim<-any(grepl("fastqc_post_trim_", rownames(files)))

hasSTAR<-any(grepl("STAR_", rownames(files)))
hasFeatureCount<-any(grepl("featureCounts_", rownames(files)))
hasDEseq2<-any(grepl("deseq2_", rownames(files)))
hasWebGestaltAnalysis<-any(grepl("WebGestalt_", rownames(files)))
hasGseaAnalysis<-any(grepl("gsea_", rownames(files)))
hasEnrichmentAnalysis<-hasWebGestaltAnalysis | hasGseaAnalysis

corNames<-gsub("_correlation_density", "", rownames(files)[grepl("_correlation_density", rownames(files))])
corNames<-sort(corNames)

getFigure<-function(category, description){
  return(paste0("```{r,echo=FALSE,results='asis', fig.align='center', fig.cap=figRef('", category, "', '",gsub("_", " ", description), "', trunk.eval=file.exists(files['", category, "',1]))}\n",
"  check_and_include_graphics(files['", category, "',1])\n```\n"))
}

if(rnaseqOptions["top25cv_in_hca", 1] == "FALSE"){
  hca_gene_used<-" using top 25% variant genes"
}else{
  hca_gene_used<-" using all genes"
}

corrFile<-"corr.Rmd"
figureRmd<-function(corNames, section){
  result<-""
  for(i in c(1:length(corNames))){
    name=corNames[i]
    heatmap<-paste0(name, "_correlation_heatmap")
    pca<-paste0(name, "_correlation_PCA")
    corr_heatmap<-paste0(name, "_correlation_cluster")
    lines<-paste0("\n\n", section, " ", name, " samples\n\n", getFigure(heatmap, paste0(heatmap, hca_gene_used)), getFigure(pca, pca), getFigure(corr_heatmap, corr_heatmap) )
    result<-paste0(result, lines)
  }
  return(result)
}
cat(figureRmd(corNames, "###"), file=corrFile)

```

<br>
  
# Sequencing Quality
  
## Summary of sequencing quality
  
```{r fastqc_raw_per_base_sequence_quality,echo=FALSE,results='asis', include = hasFastqc, eval = hasFastqc, fig.align="center", fig.cap=figRef("fastqc_raw_per_base_sequence_quality", "The mean quality scores of sequencing reads in each position", trunk.eval=file.exists(files["fastqc_raw_per_base_sequence_quality",1]))}
check_and_include_graphics(files["fastqc_raw_per_base_sequence_quality",1])
```

<br>

```{r fastqc_raw_per_base_sequence_quality_table,echo=FALSE,results='asis', include = hasFastqc, eval = hasFastqc,warning=FALSE}

tableFile=gsub(".png$","",files["fastqc_raw_per_base_sequence_quality",1])
table<-read.delim(tableFile)

QualityMin=30
if (any(table$Mean<=QualityMin)) { #use this as cutoff to show the quality table
  output_table1=data.frame(table[,c("File","Mean","Base")])
  output_table1$Base=factor(as.character(output_table1$Base),levels = unique(output_table1$Base))
  output_table1=tidyr::pivot_wider(output_table1,names_from=Base,values_from=Mean)
  output_table1=output_table1[which(apply(output_table1[,-1],1,min)< QualityMin),]
  
  print(kable(output_table1, caption=tabRef("Quality Summary", "Samples with Low RNA-seq Quality"), row.names=F) %>% 
          kable_styling() %>%
          row_spec(which(apply(output_table1[,-1],1,min)< QualityMin), color = "black", background = "honeydew") %>%
          htmltools::HTML())
}


```

<br>

```{r fastqc_raw_per_sequence_gc_content,echo=FALSE,results='asis', include = hasFastqc, eval = hasFastqc, fig.align="center", fig.cap=figRef("fastqc_raw_per_sequence_gc_content", "The average GC content of sequencing reads", trunk.eval=file.exists(files["fastqc_raw_per_sequence_gc_content",1]))}
check_and_include_graphics(files["fastqc_raw_per_sequence_gc_content",1])
```


<br>

```{r fastqc_raw_per_sequence_gc_content_table,echo=FALSE,results='hide', include = hasFastqc, eval = FALSE,warning=FALSE}

tableFile=gsub(".png$","",files["fastqc_raw_per_sequence_gc_content",1])
tableContent<-read.delim(tableFile)

tableContentSummary=tableContent %>% group_by(File) %>% summarise(GC25=sum(Count[which(GC.Content<=25)])/sum(Count),
                                              GC50=sum(Count[which(GC.Content<=50)])/sum(Count),
                                              GC75=sum(Count[which(GC.Content>=75)])/sum(Count))



GcDistributionCutoff=0.4 #GC content cutoff
gcSampleInd=which(tableContentSummary$GC25>=GcDistributionCutoff | tableContentSummary$GC75>=GcDistributionCutoff )
if (any(gcSampleInd)) { #use this as cutoff to show GC distribution
  output_table1=tableContentSummary[gcSampleInd,]
  colnames(output_table1)[-1]=c("Percent of Reads <=25% GC","Percent of Reads <=50% GC","Percent of Reads >=75% GC")
  for (i in 2:4) {
    output_table1[[i]]=paste0(round(output_table1[[i]]*100,1),"%")
  }
  
  print(kable(output_table1, caption=tabRef("GC content Distribution Summary", "Samples with abnormal GC content Distribution"), row.names=F) %>% 
          kable_styling() %>%
          htmltools::HTML() %>%
          row_spec(which(apply(output_table1[,-1],1,min)< QualityMin), color = "black", background = "honeydew"))
}


```

<br>
  
```{r fastqc_raw_adapter_content,echo=FALSE,results='asis', include = hasFastqc, eval = hasFastqc, fig.align="center", fig.cap=figRef("fastqc_raw_adapter_content", "Adapter content of sequencing reads", trunk.eval=file.exists(files["fastqc_raw_adapter_content",1]))}
check_and_include_graphics(files["fastqc_raw_adapter_content",1])
```

<br>
  
```{r fastqc_post_trim_adapter_content,echo=FALSE,results='asis', include = hasFastqc, eval = hasFastqc, fig.align="center", fig.cap=figRef("fastqc_post_trim_adapter_content", "Adapter content of sequencing reads after cutadapt", trunk.eval=file.exists(files["fastqc_post_trim_adapter_content",1]))}
check_and_include_graphics(files["fastqc_post_trim_adapter_content",1])
```

<br>
  
# Mapping quality
## Summary of mapping quality
```{r STAR_summary,echo=FALSE,results='asis', include = hasSTAR, eval = hasSTAR, fig.align="center", fig.cap=figRef("STAR_summary", "The statistics of RNAseq mapping results")}
include_graphics(files["STAR_summary",1])
```

<br>

```{r STAR_summary_table, echo=FALSE, results='asis', include = hasSTAR, eval = hasSTAR}
table<-read.csv(files["STAR_summary_table",1], check.names=F)
output_table1<-data.frame(table[,c(1,2,3,4,5)], table[,3]/table[,2], (table[,3]+table[,4])/table[,2])
colnames(output_table1)<-c("Sample", "Input reads", "Uniquely mapped reads", "Mapped to multiple loci", "Mapped to too many loci", "Uniquely mapped ratio", "Mapped ratio")

box_stats<-boxplot(output_table1$`Uniquely mapped ratio`, plot=FALSE)$stats
print(kable(output_table1, caption=tabRef("mapSummary", "The summary of RNAseq mapping results"), row.names=F) %>% 
          kable_styling() %>%
          row_spec(which(output_table1$`Uniquely mapped ratio` > box_stats[5]), color = "black", background = "bisque") %>% 
          row_spec(which(output_table1$`Uniquely mapped ratio` < box_stats[1]), color = "black", background = "honeydew") %>%
          htmltools::HTML())
cat("**Uniquely mapped ratio** = Uniquely mapped / Input reads\n\n**Mapped ratio** = (Uniquely mapped + Mapped to multiple loci)/ Input reads\n\n")
```

<br>

# Expression Quantification
## Summary of reads assignment

```{r, results="asis", echo=FALSE, include = hasFeatureCount, eval = hasFeatureCount} 
cat("featureCounts was used to count reads mapped genes, and **MULTI** mapped reads were ")
if(rnaseqOptions["featureCounts_UseMultiMappingReads", 1] == "FALSE"){
  cat(" **NOT USED**. \n\n")
}else{
  cat(" **USED**. \n\n")
}
```

```{r,echo=FALSE,results='asis', include = hasFeatureCount, eval = hasFeatureCount, fig.align="center", fig.cap=figRef("gene_reads", "The number of reads assigned to genes in uniquely mapped reads")}
# For featureCounts, unassigned_unmapped -> 
#                    unassigned_mappingquality -> 
#                    unassigned_fragmentlength -> 
#                    unassigned_chimericreads -> 
#                    unassigned_duplicate -> 
#                    unassigned_multimapping ->
#                    unassigned_secondary ->
#                    unassigned_junction_condition ->
#                    assigned
# So, we cannot get how many unique mapped reads from the featureCounts summary.
include_graphics(files["featureCounts_table_png",1])
```

<br>

```{r echo=FALSE,results='asis', include = hasFeatureCount, eval = hasFeatureCount}
table<-read.csv(files["featureCounts_table",1], check.names=F)
readsUsed = "reads"
if(rnaseqOptions["featureCounts_UseMultiMappingReads", 1] == "FALSE"){
  readsUsed = "**UNIQUE** reads"
}
colnames(table)[1]<-"Sample"
box_stats<-boxplot(table$Percent_Assigned, plot=FALSE)$stats
print(kable(table, caption=tabRef("geneSummary", paste0("The summary of ", readsUsed, " assignment to genes")), row.names=F)%>% 
          kable_styling() %>%
          row_spec(which(table$Percent_Assigned > box_stats[5]), color = "black", background = "bisque") %>% 
          row_spec(which(table$Percent_Assigned < box_stats[1]), color = "black", background = "honeydew") %>%
          htmltools::HTML())
```

<br>

## Expression density of samples
```{r correlation_density,echo=FALSE,results='asis', fig.align="center", fig.cap=figRef("correlation_density", "Expression density distribution of each sample", trunk.eval=file.exists(files["all_correlation_density",1]))}
check_and_include_graphics(files["all_correlation_density",1])
```

<br>

## Similarity between samples
```{r, child="corr.Rmd"} 
```

<br>

```{r deseq2_1, results="asis", echo=FALSE} 
if(hasDEseq2){
  cat("# Differential expression\n<br>\n\n")
  if ("DE_use_raw_pvalue" %in% row.names(rnaseqOptions) & rnaseqOptions["DE_use_raw_pvalue", 1]==1) {
    pValueText="p value"
  } else {
    pValueText="FDR adjusted p value"
  }
  deseq2criteria<-paste0("absolute fold change >= ", rnaseqOptions["DE_fold_change", 1], " and ", pValueText, " <= ", rnaseqOptions["DE_pvalue", 1])  
  cat("Differential expression analysis criteria:", deseq2criteria, "\n\n")
  cat("## Volcano plot\n<br>\n\n")
}
```

```{r deseq2_1_volcano_plot,echo=FALSE,results='asis', fig.align="center", fig.cap=figRef("Volcano_plot", "Volcano plot for all comparisons", trunk.eval=file.exists(files["deseq2_volcano_plot",1]))}
check_and_include_graphics(files["deseq2_volcano_plot",1])
```

<br>

```{r deseq2_2, results="asis", echo=FALSE} 
if(hasDEseq2 ){
  cat("## Summary of differential analysis\n<br>\n")
}
```


```{r deseq2_2_sig, echo=FALSE,results='asis'}
if(hasDEseq2 ){
  DESeq2_files<-files[grep("DESeq2_sig", files[,1]),]
  DESeq2_files<-DESeq2_files[order(DESeq2_files$V2),]
  DESeq2_list<-list()
  for (i in 1:nrow(DESeq2_files)){
    DESeq2_list[[i]]<-read.csv(DESeq2_files[i,1], header=TRUE, row.names=1, check.names=F)
  }
  diffGeneCount=sapply(DESeq2_list, function(x) 
    if("log2FoldChange" %in% colnames(x)) {
      return(c(nrow(x),length(which(x[,"log2FoldChange"]>=0)),length(which(x[,"log2FoldChange"]<0))))
    } else if ("FoldChange" %in% colnames(x)) {
      return(c(nrow(x),length(which(x[,"FoldChange"]>=1)),length(which(x[,"FoldChange"]<1))))
    } else {
      return(nrow(x))
    }
  )
#  output_table3<-data.frame(c(gsub("_DESeq2_sig.csv","",basename(DESeq2_files[,1]))), c(unlist(lapply(DESeq2_list, nrow))))
#  colnames(output_table3)<-c("Comparison", "Number of DEGs")
  output_table3<-data.frame(c(gsub("_DESeq2_sig.csv","",basename(DESeq2_files[,1]))), t(diffGeneCount))
  if (ncol(output_table3)==2) {
    colnames(output_table3)<-c("Comparison", "Number of DEGs")
  } else {
    colnames(output_table3)<-c("Comparison", "Number of DEGs","Up-regulated DEGs","Down-regulated DEGs")
  }
  print(kable(output_table3, caption= tabRef("deseq2_table", paste0("DEG numbers under FC > ", rnaseqOptions["DE_fold_change", 1], " & ",pValueText," < ", rnaseqOptions["DE_pvalue", 1]))) %>%
     kable_styling() %>%
     htmltools::HTML())

  for (i in 1:nrow(DESeq2_files)){
    comparisonName<-DESeq2_files[i,2]
    deseq2<-DESeq2_list[[i]]
    
    if(nrow(deseq2) > 0){
      maxrow<-min(nrow(deseq2), 10)
      
      cnames<-c("log2FoldChange","lfcSE", "stat", "pvalue", "padj", "FoldChange")
      hasGroupBaseMean=grepl("baseMean_", colnames(deseq2))
      if(any(hasGroupBaseMean)){
        cnames<-c(colnames(deseq2)[hasGroupBaseMean], cnames)
      }
      
      if("Feature_gene_name"%in%colnames(deseq2)){
        cnames<-c("Feature_gene_name", cnames)
      }
      deseq2table<-deseq2[c(1:maxrow), cnames]
      print(kable(deseq2table, row.names=1, caption= tabRef(comparisonName, paste0(comparisonName, " top ", maxrow, " differential expressed genes")))%>% 
          kable_styling() %>%
          row_spec(which(deseq2table$FoldChange > 1), color = "black", background = "bisque") %>% 
          row_spec(which(deseq2table$FoldChange < 1), color = "black", background = "honeydew") %>%
          htmltools::HTML())
    }
  }
}
```

<br>

```{r enrich, results="asis", echo=FALSE} 
if(hasEnrichmentAnalysis ){
  cat("# Functional enrichment analysis\nList of the top five significantly enriched elements\n<br>\n")
}
```

<br>
  
```{r webgestalt, results="asis", echo=FALSE, warning=FALSE} 
if(hasWebGestaltAnalysis){
  display_webgestalt(files)
}
```

<br>
  
```{r gsea, results="asis", echo=FALSE, warning=FALSE} 
if(hasGseaAnalysis){
  display_gsea(files, paste0(taskName, "/gsea/"))
}
``` 

<br>
  
# Results folder structure

```{r structure_link, echo=FALSE, results='asis'}

df<-NULL
countFile<-list.files('.',pattern=".count$",full.names=TRUE,recursive=TRUE)
if(length(countFile) > 0){
  df<-rbind(df, data.frame(File=addLinkTag(countFile, countFile), Description="Gene expression read count table"))
}

fpkmFile<-list.files('.',pattern="fpkm.tsv$",full.names=TRUE,recursive=TRUE)
if(length(fpkmFile) > 0){
  df<-rbind(df, data.frame(File=addLinkTag(fpkmFile, fpkmFile), Description="Gene expression abundance table"))
}

deseqAll<-list.files('.',pattern="_DESeq2.csv$",full.names=TRUE,recursive=TRUE)
if(length(deseqAll) > 0){
  df<-rbind(df, data.frame(File=addLinkTag(deseqAll, deseqAll), Description="Differential expression analysis table"))
}

deseqSig<-list.files('.',pattern="_DESeq2_sig.csv$",full.names=TRUE,recursive=TRUE)
if(length(deseqSig) > 0){
  df<-rbind(df, data.frame(File=addLinkTag(deseqSig, deseqSig), Description="Significantly differential expressed genes"))
}

webgestaltDeseq2<-list.files('.',pattern=".txt.html$",full.names=TRUE,recursive=TRUE)
if(length(webgestaltDeseq2) > 0){
  df<-rbind(df, data.frame(File=addLinkTag(webgestaltDeseq2, webgestaltDeseq2), Description="WebgestaltR enrichment table"))
}

print(kable(df, caption=tabRef("resultFiles", "Result files")) %>%
    kable_styling() %>%
    htmltools::HTML())

```

# Software versions
```{r softwareVersion, echo=FALSE, results='asis'}
versions<-display_versions()
```

# Description
```{r echo=FALSE}
vmap<-split(versions$Version, versions$Software)
has_cutadapt<-"Cutadapt" %in% names(vmap)
```

Reads were `r if(has_cutadapt){paste0("trimmed to remove adapter sequences using Cutadapt ", vmap['Cutadapt'], ")[1] and")}` aligned to the `r vmap['genome']` genome using STAR (`r vmap['STAR']`)[2]. 
`r vmap['gtf']` gene annotations were provided to STAR to improve the accuracy of mapping. 
Quality control on `r if(has_cutadapt){"both raw reads and adaptor-trimmed"}` reads was performed using FastQC (`r vmap['FastQC']`)(www.bioinformatics.babraham.ac.uk/projects/fastqc). 
featureCounts (`r vmap['featureCounts']`)[3] was used to count the number of mapped reads to each gene. 
Heatmap3 [4] was used for cluster analysis and visualization. 
`r if(hasDEseq2){paste0("Significantly differential expressed genes with ", deseq2criteria, " were detected by DESeq2 (", vmap['DESeq2'], ")[5].")}`
`r if(hasWebGestaltAnalysis){paste0("Genome Ontology and KEGG pathway over-representation analysis was performed on differentially expressed genes using the WebGestaltR package (", vmap['WebGestaltR'], ")[6].")}`
`r if(hasGseaAnalysis){paste0("Gene set enrichment analysis was performed using GSEA package (", vmap['GSEA'], ")[7] on database (", vmap['GSEA_DB'], ").")}`

`r if(has_cutadapt){"[1]	Martin M: Cutadapt removes adapter sequences from high-throughput sequencing reads. EMBnetjournal 2011, 17:10-12."}` \
[2]	Dobin A, Davis CA, Schlesinger F, Drenkow J, Zaleski C, Jha S, Batut P, Chaisson M, Gingeras TR: STAR: ultrafast universal RNA-seq aligner. Bioinformatics 2013, 29(1):15-21. \
[3]	Liao Y, Smyth GK, Shi W: featureCounts: an efficient general purpose program for assigning sequence reads to genomic features. Bioinformatics 2014, 30(7):923-930. \
[4]	Zhao S, Guo Y, Sheng Q, Shyr Y: Advanced heat map and clustering analysis using heatmap3. BioMed research international 2014, 2014:986048. \
`r if(hasDEseq2){"[5] Love MI, Huber W, Anders S: Moderated estimation of fold change and dispersion for RNA-seq data with DESeq2. Genome Biol 2014, 15(12):550."}` \
`r if(hasWebGestaltAnalysis){"[6]	Wang J, Vasaikar S, Shi Z, Greer M, Zhang B: WebGestalt 2017: a more comprehensive, powerful, flexible and interactive gene set enrichment analysis toolkit. Nucleic Acids Res 2017, 45(W1):W130-W137."}` \
`r if(hasGseaAnalysis){"[7]	Subramanian A, Tamayo P, Mootha VK, Mukherjee S, Ebert BL, Gillette MA, Paulovich A, Pomeroy SL, Golub TR, Lander ES et al: Gene set enrichment analysis: a knowledge-based approach for interpreting genome-wide expression profiles. Proc Natl Acad Sci U S A 2005, 102(43):15545-15550."}` \
