---
title: "linear association analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: hide
    number_sections: true
---

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo=TRUE, 
  include=TRUE, 
  warning=FALSE, 
  message=FALSE, 
  error=FALSE,
  results="asis"
)
```

```{r}
source('reportFunctions.R')
library(arsenal)
library(xtable)

option_tbl=fread('input_options.txt', header=FALSE)
cur_options=split(option_tbl$V1, option_tbl$V2)

phename=cur_options$phename
phecode=as.numeric(cur_options$phecode)
phefile=cur_options$phefile
if (file.exists("../fileList2.list")){
  global_tbl=fread("../fileList2.list", header=FALSE)
  global_options=split(global_tbl$V1, global_tbl$V2)
}else{
  global_options=cur_options
}

agd_file=global_options$agd_file
demographics_file=global_options$demographics_file
genotype_file=global_options$genotype_file
pca_file=global_options$pca_file
phecode_map_file=global_options$phecode_map_file
ancestry_file=global_options$ancestry_file

email=global_options$email
affiliation=global_options$affiliation

if(!is.null(email)){
  if(email != ""){
    cat("
  ---
  author: 
  - name: ", email, "
    affiliation: ", affiliation, "
  ---
  ")
  }
}
```

# Introduction

This document is used to perform correlation analysis between genotype and phenotype using AGD 35K cohort.

# Read phenotype data

Read data from `r phefile`.

```{r}
pheno_data=fread(phefile, data.table=FALSE)
print_table(head(pheno_data), row.names=FALSE)
```

# Read genotype data

Read data from `r genotype_file`.

```{r}
geno_data=fread(genotype_file, data.table=FALSE)
print_table(head(geno_data), row.names=FALSE)
```

# Read PCA data

Read data from `r pca_file`.

```{r}
pca_data=fread(pca_file, data.table=FALSE) |>
  dplyr::rename("GRID"="IID") |>
  dplyr::select(-`#FID`)
print_table(head(pca_data), row.names=FALSE)
```

# Read Race data

```{r}
ancestry_data=fread(ancestry_file, data.table=FALSE) |>
  dplyr::mutate(Age=2023 - YEAR_OF_BIRTH,
                Gender=GENDER_SOURCE_VALUE,
                Race=replace(supervised_ancestry_cluster, supervised_ancestry_cluster=="Admixed (majority ancestry < 0.5)", "Admixed")) |>
  dplyr::select(GRID, Age, Gender, Race)
print_table(head(ancestry_data), row.names=FALSE)
```

# Combine data

```{r}
combined_data=merge(pheno_data, geno_data, by='GRID' ,all.x=TRUE) |>
  merge(pca_data, by='GRID', all.x=TRUE) |>
  merge(ancestry_data, by='GRID', all.x=TRUE)

combined_data$Phenotype=as.factor(combined_data$Phenotype)
combined_data$Genotype=as.factor(combined_data$Genotype)
combined_data$Gender=as.factor(combined_data$Gender)
combined_data$Race=as.factor(combined_data$Race)

print_table(combined_data, byDT=TRUE, row.names=FALSE)  
```

# Summary of data grouping by Phenotype

```{r}
mycontrols  <- tableby.control(test=T, total=T,
                               numeric.stats=c("Nmiss","median","q1q3","range"),
                               cat.stats=c("Nmiss","countpct"),
                               stats.labels=list(Nmiss='Missing', median='Median', range='Range'),
                               digits=1, digits.p=2, digits.pct=0)

table_sum1 <- tableby(Phenotype ~ Genotype + Age + Gender + Race, data = combined_data, control=mycontrols)
summary(table_sum1, total=T, title = "", width = 3, digits=1, digits.p=2, digits.pct=0, pfootnote = TRUE)
```

# Correlation analysis

```{r}
res <- glm(Phenotype ~ Genotype + Age + Gender + Race + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, family=binomial(link="logit"), data=combined_data)
coef=summary(res)$coefficients |>
  as.data.frame() 
rownames(coef)[2]="Genotype"
print(xtable(coef, digits=-2, caption="Logistic regression coefficients"),type='html')
```

```{r}
coef_file=paste0(phename, ".linear_association.csv")
write.csv(coef, coef_file, quote=FALSE, row.names=TRUE)
```

coefficients are saved to `r coef_file`.
