---
title: "Phenotype data prepration"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: hide
    number_sections: true
---

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo=TRUE, 
  include=TRUE, 
  warning=FALSE, 
  message=FALSE, 
  error=FALSE,
  results="asis"
)
```


```{r}
source('reportFunctions.R')

option_tbl=fread('input_options.txt', header=FALSE)
cur_options=split(option_tbl$V1, option_tbl$V2)

phename=cur_options$phename
phecode=as.numeric(cur_options$phecode)
option_file=cur_options$option_file
global_tbl=fread(option_file, header=FALSE)
global_options=split(global_tbl$V1, global_tbl$V2)

agd_file=global_options$agd_file
phecode_data_file=global_options$phecode_data_file
phecode_map_file=global_options$phecode_map_file
min_occurance=as.numeric(global_options$min_occurance)
email=global_options$email
affiliation=global_options$affiliation

if(email != ""){
  cat("
---
author: 
- name: ", email, "
  affiliation: ", affiliation, "
---
")
}
```

# Introduction

This document is used to prepare phenotype data for GWAS/PheWAS analysis using AGD cohort.

phecode definition file was downloaded from `https://phewascatalog.org/phewas/_w_e1b53007/data/phecode_definitions1.2.csv.zip` at Nov 1st, 2024. 

# Read phecode data from AGD cohort.

Read data from `r phecode_data_file`.

```{r}
phecodes=fread(phecode_data_file, data.table=FALSE)
print_table(head(phecodes), row.names=FALSE)
```

# Read phecode definition file

Read data from `r phecode_map_file`.

```{r}
phecodes_def=fread(phecode_map_file, data.table=FALSE) |>
  dplyr::mutate(exclude_start=as.numeric(gsub("\\-.+", "", phecode_exclude_range)),
                exclude_end=as.numeric(gsub(".+\\-", "", phecode_exclude_range)),
                phecode_str=phenotype) |>
  dplyr::select(phecode, phecode_str, exclude_start, exclude_end)
print_table(head(phecodes_def), row.names=FALSE)
```

# Filter `r phename` definition

We extracted `r phename` definition using phecode `r phecode`.

```{r}
cur_def=phecodes_def[phecodes_def$phecode==phecode,] |>
  dplyr::select(phecode, phecode_str, exclude_start, exclude_end) |>
  dplyr::distinct()

exclude_start=cur_def$exclude_start[1]
exclude_end=cur_def$exclude_end[1]
phecode_str=cur_def$phecode_str[1]

print_table(cur_def, row.names=FALSE)
```

# Find positive subjects

`r phecode_str` subjects would have phecode `r phecode`.

```{r}
phecode_samples = phecodes[phecodes$phecode==phecode,] |>
    dplyr::group_by(GRID) |>
    dplyr::summarize(n=n())

if(min_occurance > 1){
  cat("Only subjects with at least ", min_occurance, " occurances are considered.\n")
  pos_samples = phecode_samples |>
    dplyr::filter(n >= min_occurance) 
}else{
  pos_samples = phecode_samples
}
print_table(pos_samples, byDT=TRUE, row.names=FALSE)
```

There are `r nrow(pos_samples)` positive subjects.

# Find negative subjects

Read data from `r agd_file`.

```{r}
AGD=fread(agd_file, data.table=FALSE) |>
  dplyr::filter(PRIMARY_GRID != "-")
```

There are `r nrow(AGD)` eligible subjects in AGD cohort.

Let's find possible positive subjects with phecode between `r exclude_start` and `r exclude_end`.

```{r}
possible_pos_samples = phecodes |>
  dplyr::filter(phecode >= exclude_start) |>
  dplyr::filter(phecode <= exclude_end) |>
  dplyr::distinct() |>
  dplyr::arrange(GRID)
print_table(possible_pos_samples, byDT=TRUE, row.names=FALSE)
```

There are `r length(unique(possible_pos_samples$GRID))` possible positive subjects.

Let's find negative subjects in AGD cohort.

```{r}
negative_samples = AGD |>
  dplyr::filter(!PRIMARY_GRID %in% possible_pos_samples$GRID)
```

There are `r nrow(negative_samples)` negative subjects.

# Save phenotype data

```{r}
cur_phenotype_def = data.frame( GRID=c(pos_samples$GRID, negative_samples$PRIMARY_GRID),
                                Phenotype=rep(c("1", "0"), c(nrow(pos_samples), nrow(negative_samples))))

phenotype_file = paste0(phename, '.phenotype.csv')
write.csv(cur_phenotype_def, phenotype_file, quote=FALSE, row.names=FALSE)
print_table(cur_phenotype_def, byDT=TRUE, row.names=FALSE)
```

We have `r nrow(pos_samples)` positive subjects and `r nrow(negative_samples)` negative subjects.

Phenotype information for `r phecode_str` is saved to file [`r phenotype_file`](`r phenotype_file`).
