---
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output: 
  rmdformats::readthedown:
    toc_depth: 4
    code_folding: hide
    number_sections: yes
---

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, echo = F, message=FALSE}
knitr::opts_chunk$set(
  echo=T,  
  message=FALSE, 
  warning=FALSE, 
  results = 'asis', 
  fig.width=7, 
  fig.height = 7
)
```

```{r}
library(ggsci)
library(ggplot2)
library(pheatmap)
library(limma)
library(DT)
library(reshape2)
library(data.table)
library(htmltools)
library(randomcoloR)

source('reportFunctions.R')
source('lipidomics_func.R')
source('countTableVisFunctions.R')

myoptions=read_file_map("fileList1.limma.txt", do_unlist=FALSE)
email=myoptions$email
task_name=myoptions$task_name
affiliation=myoptions$affiliation
heatmap_width=as.numeric(myoptions$heatmap_width)

foldchange <- as.numeric(myoptions$fold_change)
fdr <- as.numeric(myoptions$fdr)

if(!"remove_sample_pattern" %in% names(myoptions)){
  myoptions$remove_sample_pattern <- ""
}
if(!"remove_sample_description" %in% names(myoptions)){
  myoptions$remove_sample_description <- ""
}
```

---
title: Differential analysis with Limma of Lipidomics Project `r gsub(":", " ", task_name)`
author:
- name: `r email`
  affiliation: `r affiliation`
---

# Input files

## Metadata 

```{r}
meta_file <- fread("fileList3.limma.txt", header=F, data.table = F)$V1[1]
```

Read `r meta_file` file

```{r}
meta <- fread(meta_file, data.table = F)
sample_group_map=unlist(split(meta$Class, meta$Class_Sample))
datatable(meta, rownames = F)
```

## Data files

```{r}
list_file <- fread("fileList2.limma.txt", header=F, data.table = F)$V1[1]
```

Read `r list_file` file

```{r}
files <- fread(list_file, header = T, data.table = F)
datatable(files, rownames = F)
```

Read data files, rename Sample with Class_Sample.

```{r}
raw_files_list <- read_data_files(files, "RawFile")
log_files_list <- read_data_files(files, "LogFile")
```

```{r}
remove_samples<-function(files_list, pattern){
  result <- lapply(files_list, function(x){
    y <- x[, -grep(pattern, colnames(x))]
    return(y)
  })
  return(result)
}

if(myoptions$remove_sample_pattern != ""){
  cat("\n# Remove samples\n\n")
  cat(myoptions$remove_sample_description, "\n\n")

  raw_files_list_sel <- remove_samples(files_list=raw_files_list, pattern=myoptions$remove_sample_pattern)
  log_files_list_sel <- remove_samples(files_list=log_files_list, pattern=myoptions$remove_sample_pattern)

  meta_sel <- meta[-grep(myoptions$remove_sample_pattern, meta$Class_Sample),]
  datatable(meta_sel, rownames = F)
}else{
  raw_files_list_sel <- raw_files_list
  log_files_list_sel <- log_files_list

  meta_sel=meta
}
rm(raw_files_list, log_files_list, meta)
```

```{r}
png_list_file <- fread("fileList5.limma.txt", header=F, data.table = F)$V1[1]
png_files <- data.frame(fread(png_list_file, header = T), row.names=1)
```

# Comparisons

```{r}
comparison_file <- fread("fileList4.limma.txt", header=F, data.table = F)
comp_table <- data.frame(Comparison = unique(comparison_file$V2), Reference="", Sample="")

idx=1
for(idx in 1:nrow(comp_table)){
  cur_comp=comp_table$Comparison[idx]
  cur_tbl=comparison_file[comparison_file$V2==cur_comp,]
  comp_table[idx, "Reference"] <- cur_tbl[1,1]
  comp_table[idx, "Sample"] <- cur_tbl[2,1]
}

comp_table$Contract = paste(comp_table$Sample, "-", comp_table$Reference)

print_table(comp_table, row.names=FALSE)
```

# Differential expression analysis

Fold change > `r foldchange` and FDR < `r fdr ` are considered significant   

```{r}
#limma contrasts
design <- model.matrix(~0 + factor(meta_sel$Class, levels = unique(meta_sel$Class)))
rownames(design) <- meta_sel$Sample
colnames(design) <- unique(meta_sel$Class)

contrasts_tomake=comp_table$Contract
contrast.matrix <- eval(as.call(c(as.symbol("makeContrasts"),as.list(contrasts_tomake),levels=list(design))))
```

```{r}
#limma
de_list_bay <- list()
for (i in 1:length(log_files_list_sel)){
  fit <- lmFit(log_files_list_sel[[i]], design)
  fitC <- contrasts.fit(fit, contrast.matrix)
  de_list_bay[[i]] <- eBayes(fitC)
}
```

```{r, results='asis'}
# All results
de_list_toptable <- list()
for (i in 1:length(de_list_bay)){
    de_list_toptable[[i]] <- list()
    for (j in 1:length(contrasts_tomake)){
        de_list_toptable[[i]][[j]] <- topTable(de_list_bay[[i]], coef = contrasts_tomake[j], adjust = "fdr", sort.by = "t", number=nrow(log_files_list_sel[[i]]))
    }
    names(de_list_toptable[[i]]) <- contrasts_tomake
}
names(de_list_toptable) <- names(log_files_list_sel)

#Significant results
de_list_toptable_sig <- list()
for (i in 1:length(de_list_bay)){
    de_list_toptable_sig[[i]] <- list()
    for (j in 1:length(contrasts_tomake)){
        de_list_toptable_sig[[i]][[j]] <- topTable(de_list_bay[[i]], coef = contrasts_tomake[j], adjust = "fdr", sort.by = "t", number=nrow(log_files_list_sel[[i]]), p.value = fdr, fc = foldchange)
    }
    names(de_list_toptable_sig[[i]]) <- contrasts_tomake
}
names(de_list_toptable_sig) <- names(log_files_list_sel)
```

```{r}
#plotting functions
## Heatmap
heatmap_heights=NA

## PCA
pca_width=6
pca_height=5

# Volcano plot
volcano_width=7
volcano_height=6

#for volcano plot
reverselog_trans <- function(base = exp(1)) {
  trans <- function(x) -log(x, base)
  inv <- function(x) base^(-x)
  trans_new(paste0("reverselog-", format(base)), trans, inv, 
            log_breaks(base = base), 
            domain = c(1e-100, Inf))
}

library(EnhancedVolcano)

draw_volcano <- function(data, category, comparison, prefix, volcano_width, volcano_height, fdr, foldchange){
  diffResult<-as.data.frame(data)
  yname=bquote(-log[10](p~value))
  p<-EnhancedVolcano(diffResult,
      lab = rownames(diffResult),
      x = 'logFC',
      y = 'P.Value',
      title = comparison,
      pCutoff = fdr,
      pCutoffCol = "adj.P.Val",
      FCcutoff = log2(foldchange),
      pointSize = 3.0,
      labSize = 3.0,
      colAlpha = 1,
      subtitle = NULL) + ylab(yname)

  png_file = paste0(prefix, ".volcano.png")
  ggsave(png_file, plot = p, width = volcano_width, height = volcano_height, dpi = 300, units = "in", bg="white")
  return(png_file)
}

```

```{r, results='asis'}

#loop printing tables
#for (i in 1:length(de_list_toptable_sig)){
#  for(j in 1:length(de_list_toptable_sig[[i]])){
#  #Save results
#  write.csv(de_list_toptable[[i]][[j]], paste(task_name, names(de_list_toptable)[[i]], names(de_list_toptable[[i]])[[j]], "DE.csv", sep = "_"))
#    if(nrow(de_list_toptable_sig[[i]][[j]]) == 0){next}
#        write.csv(de_list_toptable_sig[[i]][[j]], paste(task_name, names(de_list_toptable_sig)[[i]], names(de_list_toptable_sig[[i]])[[j]], "DE_sig.csv", sep = "_"))
#        cat("\n") 
#        cat("## ", names(de_list_toptable_sig)[i], "\n")
#        cat("### ", names(de_list_toptable_sig[[i]])[j], "\n")
#        print(tagList(datatable(de_list_toptable_sig[[i]][[j]])))
#        cat("\n")
#  }
#}


```

```{r, results='asis', fig.height = 5, fig.width = 5}
#loop printing tables, pca, heatmap, volcano plot
groups <- data.frame(row.names = meta_sel$Class_Sample, Group = meta_sel$Class)

pvalue<-fdr
useRawPvalue<-0

cur_folder = getwd()
limma_folder = paste0(cur_folder, "/limma_result/")
if(!dir.exists(limma_folder)){
  dir.create(limma_folder)
}

de_tbl = data.frame("Category"=character(), 
                    "Comparison"=character(), 
                    "Total_Feature"=numeric(), 
                    "DE_feature"=numeric(), stringsAsFactors=FALSE)
result = ""
i=1
de_idx = 0
for (i in 1:length(de_list_toptable_sig)){
  category = names(de_list_toptable_sig)[i]
  category_df_list = de_list_toptable[[i]]
  category_sig_df_list = de_list_toptable_sig[[i]]
  category_log_df = log_files_list_sel[[i]]

  result = paste0(result, "\n## ", category, "\n\n")

  result = paste0(result, "\n### Global PCA\n\n")
  pca_png = png_files[category, "pca_file"]
  result = paste0(result, getFigure(pca_png))

  j=1
  for(j in 1:length(category_sig_df_list)){
    de_idx = de_idx + 1

    contrast_name = names(category_sig_df_list)[[j]]
    cur_comp_tbl=comp_table[comp_table$Contract==contrast_name,]
    cur_comp=cur_comp_tbl[1,"Comparison"]
    cur_reference_group=cur_comp_tbl[1,"Reference"]
    cur_sample_group=cur_comp_tbl[1,"Sample"]
    cur_comp_df = category_df_list[[j]]
    cur_comp_sig_df = category_sig_df_list[[j]]

    groups_colors = c("red", "blue")
    names(groups_colors) = c(cur_sample_group, cur_reference_group)
    annotation_colors = list(Group=groups_colors)

    cur_prefix = paste0(limma_folder, paste(task_name, category, cur_comp, sep = "."))
    #Save results
    write.csv(cur_comp_df, paste0(cur_prefix, ".DE.csv"))
  
    result = paste0(result, "\n### ", contrast_name, "\n\n") 

    cur_samples = meta_sel$Class_Sample[meta_sel$Class %in% c(cur_comp_tbl[1,"Sample"], cur_comp_tbl[1,"Reference"])]
    cur_groups = groups[cur_samples,,drop=FALSE]

    data_df = category_log_df[,cur_samples]
    if(min(rowSums(data_df)) == 0) {
      data_df = data_df[-which(rowSums(data_df) == 0),]}

    show_rownames=nrow(data_df) <= 30
    if(show_rownames){
      cur_heatmap_width = heatmap_width + 2
    }else{
      cur_heatmap_width = heatmap_width
    } 

    heatmap_height = heatmap_heights[i]
    png_file = draw_pheatmap( data_df = data_df, 
                              groups = cur_groups, 
                              annotation_colors = annotation_colors, 
                              title=category, 
                              file_prefix=cur_prefix,
                              heatmap_width=cur_heatmap_width,
                              heatmap_height=heatmap_height,
                              annotation_legend=TRUE,
                              show_rownames=show_rownames)

    result = paste0(result, "\n#### Heatmap\n\n")
    if(is.na(heatmap_height)){
      result = paste0(result, getFigure_width_height(png_file, fig.width=heatmap_width, fig.height=NULL))
    }else{
      result = paste0(result, getFigure_width_height(png_file, fig.width=heatmap_width, fig.height=heatmap_height))
    }

    pca_groups = cur_groups$Group
    png_file = draw_pca(data=data_df, 
                        groups=pca_groups, 
                        title=category, 
                        file_prefix=cur_prefix, 
                        pca_width=pca_width, 
                        pca_height=pca_height,
                        colors = groups_colors[match(unique(pca_groups), names(groups_colors))]
                        )
    result = paste0(result, "\n#### PCA\n\n")
    result = paste0(result, getFigure_width_height(png_file, fig.width=pca_width, fig.height=pca_height))

    de_tbl[de_idx,] = c(category, contrast_name, nrow(cur_comp_df), nrow(cur_comp_sig_df))

    #Tables
    if(nrow(cur_comp_sig_df) == 0){
      result = paste0(result, "\nNo significant lipids found.\n\n")
      next
    }

    sig_csv = paste0(cur_prefix, ".DE.sig.csv")
    write.csv(cur_comp_sig_df, sig_csv)
      
    result = paste0(result, "\n#### DE features\n\n")
    result = paste0(result, getPagedTable(sig_csv))

    #Volcano plot
    png_file = draw_volcano(data=cur_comp_df, 
                            category=category, 
                            comparison=contrast_name,
                            prefix=cur_prefix, 
                            volcano_width=volcano_width, 
                            volcano_height=volcano_height,
                            fdr=fdr,
                            foldchange=foldchange)
    result = paste0(result, "\n#### Volcano\n\n")
    result = paste0(result, getFigure_width_height(png_file, fig.width=volcano_width, fig.height=volcano_height))
  }
}
de_rmd_file = paste0(task_name, ".de.rmd")
writeLines(result, de_rmd_file)

de_file = paste0(task_name, ".de_summary.csv")
write.csv(de_tbl, de_file)
```
```{r, child=de_rmd_file}
```

# DE analysis summary

```{r}
print_table(de_tbl, row.names=FALSE)
```
