```{r}
library("R.utils")
```

# Filter criteria

- Metabolite.name != 'Unknown' to remove unknown features
- Metabolite.name doesn't contain 'w/o' to remove features without ms2
- Ontology != 'Others' and metabolite.name doesn't contain 'RIKEN' to remove other unidentified features
- Metabolite.name doesn't contain '(d7)' or '(d9)' to remove internal standards
- Mean(QC) > 0, RSD(QC) < 25%, Mean(Blank)/Mean(QC) < 0.1 to remove low quality features

# Preprocessing positive data

Read positive area data from `r pos_file`.

```{r}
lst<-read_lipid_tsv_and_filter(filename=pos_file, filePrefix=paste0(result_prefix, ".pos"))

pos_df=lst$res
pos_all_meta=lst$all_meta
pos_sample_meta=lst$sample_meta
pos_filter_tb=lst$filter_tb
```

## Pos all meta data

```{r}
print_table(pos_all_meta, row.names=FALSE, byDT=TRUE)
```

## Pos sample meta data

```{r}
print_table(pos_sample_meta, row.names=FALSE, byDT=TRUE)
```

## Pos filter table

```{r}
print_table(pos_filter_tb, row.names=FALSE)
```

## Pos data

```{r}
print_table(head(pos_df), row.names=FALSE)
```

# Preprocessing negative data

Read negative area data from `r neg_file`.

```{r}
lst<-read_lipid_tsv_and_filter(filename=neg_file, filePrefix=paste0(result_prefix, ".neg"))

neg_df=lst$res
neg_all_meta=lst$all_meta
neg_sample_meta=lst$sample_meta
neg_filter_tb=lst$filter_tb
```

## Neg all meta data

```{r}
print_table(neg_all_meta, row.names=FALSE, byDT=TRUE)
```

## Neg sample meta data

```{r}
print_table(neg_sample_meta, row.names=FALSE, byDT=TRUE)
```

## Neg filter table

```{r}
print_table(neg_filter_tb, row.names=FALSE)
```

## Neg data

```{r}
print_table(head(neg_df), row.names=FALSE)
```

# Combine positive and negative data

```{r}
stopifnot(all(colnames(pos_df)==colnames(neg_df)))
```

Sample names in pos and neg files are identical.

We will combine positive and negative data by `r combine_by`.

- HIGHEST: use the highest value of same molecule from the two modes.
- SUM: sum the values of same molecule from the two modes.

```{r}
pos_df$Mode="POS"
neg_df$Mode="NEG"
combined=rbind(pos_df, neg_df) %>% relocate(Mode, .after=Ontology)

get_highest_value<-function(combined){
  dupnames=combined$Metabolite.name[duplicated(combined$Metabolite.name)]
  final=combined[!combined$Metabolite.name %in% dupnames,]

  #find the entry with highest value in most samples
  dupname=dupnames[1]
  for(dupname in dupnames){
    cur_dup=combined[combined$Metabolite.name==dupname,] %>% select(-Metabolite.name, -Ontology, -Mode)
    cur_rank = apply(cur_dup, 2, function(y) rank(-y))
    sum_rank=rowSums(cur_rank)
    min_rank=which(sum_rank==min(sum_rank))
    min_dup=cur_dup[min_rank,]
    final=rbind(final, combined[rownames(min_dup),])
  }  
  return(final)
}

get_sum_value<-function(combined){
  dupnames=combined$Metabolite.name[duplicated(combined$Metabolite.name)]
  final=combined[!combined$Metabolite.name %in% dupnames,]

  dupname=dupnames[1]
  for(dupname in dupnames){
    cur_dup=combined[combined$Metabolite.name==dupname,]
    # It is possible that all the duplicates are from same mode.
    # We need to check the mode instead of just assign "BOTH"
    if(length(unique(cur_dup$Mode))==2){ 
      cur_dup$Mode="BOTH"
    }

    cur_res=aggregate(. ~ Metabolite.name + Ontology + Mode, cur_dup, sum)
    final=rbind(final, cur_res[,colnames(final)])
  }  
  return(final)
}

if(combine_by == "SUM"){
  final=get_sum_value(combined)
}else if(combine_by == "HIGHEST"){
  final=get_highest_value(combined)
}

combined_file=paste0(result_prefix, ".combined.csv")
write.csv(final, combined_file, row.names=FALSE)
print_table(final, row.names=FALSE, byDT=TRUE)
```

After combining POS and NEG data, removing duplicate molecules, there are total `r nrow(final)` metabolites. 
Data was saved to <mark>`r combined_file`</mark>.

# Add annotation

```{r}
anno=fread(anno_file, data.table=FALSE)
print_table(anno, byDT=TRUE, row.names=FALSE)
```

```{r}
venn.width=4
venn.height=4
```

The Ontology in data file corresponds to Abbreviation in MSDial annotation table.

```{r}
x<-list("MSDial"=c(unique(anno$Abbreviation)), "Data"=unique(final$Ontology))
g<-ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF"),
  stroke_size = 0.5, set_name_size = 4
  )
venn_file = paste0(result_prefix, ".MSDial_data.venn.png")
ggsave(venn_file, g, width=venn.width, height=venn.height, units="in", dpi=300, bg="white")
knitr::include_graphics(venn_file)
```

```{r}
missed=final[!(final$Ontology %in% anno$Abbreviation),]
cat("\n\nThere", ifelse(nrow(missed)<2, "is", "are"), nrow(missed), "molecule(s) not found in MSDial annotation file.\n\n")
```

```{r, eval=nrow(missed)>0}
print_table(missed, row.names=FALSE)
```

# Final annotated tables

```{r}
anno_slim=anno[,c(1:4)]
anno_slim=anno_slim[!duplicated(anno_slim$Abbreviation),]

final_anno=dplyr::left_join(final, anno_slim, join_by(Ontology==Abbreviation)) %>%
  dplyr::rename(Abbreviation=Ontology) %>%
  select(c("Metabolite.name", "Mode", "Abbreviation", "Lipid_subclass","Main_class","Categories"), everything()) %>%
  rename("Lipid_subclass_abbreviation"="Abbreviation")
final_anno$Main_class[is.na(final_anno$Main_class)] = "Others"
final_anno$Categories[is.na(final_anno$Categories)] = "Others"
rownames(final_anno)<-final_anno$Metabolite.name
samples=colnames(final_anno)[7:ncol(final_anno)]

get_box<-function(bdf, samples){
  bdf=data.frame(bdf)
  bdf$feature<-rownames(bdf)
  mbdf<-melt(bdf,id.vars="feature")
  mbdf<-mbdf[order(mbdf$feature),]
  
  nf=min(10, length(unique(bdf$feature)))
  
  plot_df=mbdf[c(1:(nf * length(samples))),]
  g<-ggplot(plot_df, aes(y=feature, x=value)) + geom_boxplot() + theme_bw()
  return(g)
}

log_scale<-function(df, samples, data_prefix, box_width, box_height){
  raw_csv=paste0(data_prefix, ".raw.csv")
  write.csv(df, raw_csv)
  g0<-get_box(df, samples) + xlab("Concentration") + ggtitle("pre-log-scaling")
  
  ldf<-log10(df+1)
  log_csv=paste0(data_prefix, ".log.csv")
  write.csv(ldf, log_csv)
  g1<-get_box(ldf, samples) + xlab("Concentration") + ggtitle("post-log")

  sldf<-data.frame(t(scale(t(ldf), center=TRUE, scale=TRUE)))
  log_scaled_csv=paste0(data_prefix, ".log_scaled.csv")
  write.csv(sldf, log_scaled_csv)
  g2<-get_box(sldf, samples) + xlab("Concentration") + ggtitle("post-log-scaling")
  
  g<-g0+g1+g2+plot_layout(ncol=3)
  
  log_scaled_png=paste0(data_prefix, ".log_scaled.png")
  ggsave(log_scaled_png, width=box_width, height=box_height, dpi=300, units="in", bg="white")
  return(list(raw_csv=getAbsolutePath(raw_csv), log=ldf, log_csv=getAbsolutePath(log_csv), log_scaled=sldf, log_scaled_csv=getAbsolutePath(log_scaled_csv), log_scaled_png=log_scaled_png))
}

box_width=13
box_height=4
```

```{r}
full_file=paste0(annotated_prefix, ".full.raw.csv")
write.csv(final_anno, full_file, row.names=F)
print_table(final_anno, byDT=TRUE, row.names=F)
```

Raw concentration file was saved to `r full_file`.

## Species

```{r}
df=final_anno[,samples]
lst=log_scale(df, samples, paste0(annotated_prefix, ".species"), box_width, box_height)

final_files=data.frame("RawFile"=lst$raw_csv, "LogFile"=lst$log_csv, "Name"="Species")
```

```{r child="lipidomics_category.Rmd"}
```

## Sub class

```{r}
final_sub_class=final_anno[,c("Lipid_subclass_abbreviation",samples),drop=FALSE]

final_sub_class=aggregate(. ~ Lipid_subclass_abbreviation, final_sub_class, sum)
colnames(final_sub_class)<-c("Lipid_subclass", samples)
final_sub_class <- final_sub_class %>% tibble::column_to_rownames("Lipid_subclass")

lst=log_scale(final_sub_class, samples, paste0(annotated_prefix, ".sub_class"), box_width, box_height)
final_files=rbind(final_files, data.frame("RawFile"=lst$raw_csv, "LogFile"=lst$log_csv, "Name"="Sub_class"))
```

```{r child="lipidomics_category.Rmd"}
```

## Main class

```{r fig.width=box_width, fig.height=box_height}
final_main_class=final_anno[,c("Main_class",samples),drop=FALSE]
final_main_class=aggregate(. ~ Main_class, final_main_class, sum)
colnames(final_main_class)<-c("Main_class", samples)
final_main_class <- final_main_class %>% tibble::column_to_rownames("Main_class")

lst=log_scale(final_main_class, samples, paste0(annotated_prefix, ".main_class"), box_width, box_height)
final_files=rbind(final_files, data.frame("RawFile"=lst$raw_csv, "LogFile"=lst$log_csv, "Name"="Main_class"))
```

```{r child="lipidomics_category.Rmd"}
```

## Category

```{r fig.width=box_width, fig.height=box_height}
final_category=final_anno[,c("Categories",samples),drop=FALSE]

final_category=aggregate(. ~ Categories, final_category, sum)
colnames(final_category)=c("Categories", samples)
final_category <- final_category %>% tibble::column_to_rownames("Categories")

lst=log_scale(final_category, samples, paste0(annotated_prefix, ".category"), box_width, box_height)
final_files=rbind(final_files, data.frame("RawFile"=lst$raw_csv, "LogFile"=lst$log_csv, "Name"="Category"))
```

```{r child="lipidomics_category.Rmd"}
```

```{r}
write.table(final_files, paste0(task_name, ".files.txt"), quote=FALSE, row.names=FALSE, col.names=TRUE, sep="\t")
```
